---
title: "Aflevering"
output: pdf_document
date: "2026-05-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Library
```{r}
library(mlr3)
library(mlr3learners)
library(mlr3tuning)
library(mlr3mbo)
library(mlr3pipelines)
library(paradox)
library(readr)
library(tidyverse)
library(mlr3viz)
```

# Load data
Below is data loaded, and ajustments are made: 
```{r}
freMPL <- read_csv("freMPL.csv")

# VehAge is an ordered variable, this ordering is taken into consideration
freMPL$VehAge <- factor(freMPL$VehAge, levels = c("0", "1", "2", "3", "4", "5", "6-7", "8-9", "10+"))

# VehMaxSpeed is an ordered variable, this ordering is taken into consideration
freMPL$VehMaxSpeed <- factor(freMPL$VehMaxSpeed,
                             levels = c("1-130 km/h", "130-140 km/h", "140-150 km/h",
                                        "150-160 km/h", "160-170 km/h", "170-180 km/h", 
                                        "180-190 km/h", "190-200 km/h", "200-220 km/h",
                                        "220+ km/h"))

# Do we need to do something to SocioCateg?
# Why do ClaimInt = 0 have negative ClaimAmount? subset = freMPL[freMPL$ClaimInd == 0,]
```

```{r}
View(freMPL)
summary(freMPL)
table(freMPL$VehAge)
table(freMPL$MariStat)
table(freMPL$SocioCateg)
```

```{r}
plot(freMPL$DrivAge, freMPL$LicAge)
plot(freMPL$DrivAge, freMPL$BonusMalus)
plot(freMPL$DrivAge, freMPL$ClaimInd)
```

```{r}
table(freMPL$VehBody, freMPL$VehUsage)
```
```{r}
names(freMPL)
```

An overview of all the covariates: 
```{r}
# Exposure
ggplot(freMPL, aes(x = Exposure)) + 
  geom_histogram(bins = 50, fill = "blue", col = "white") + 
  theme_minimal()
# LicAge
ggplot(freMPL, aes(x = LicAge/12)) + 
  geom_histogram(bins = 50, fill = "blue", col = "white") + 
  xlab("LicAge in years") + 
  theme_minimal()
# RecordBeg
ggplot(freMPL, aes(x = RecordBeg)) + 
  geom_histogram(bins = 13, fill = "blue", col = "white") + 
  theme_minimal()
# RecordEnd
ggplot(freMPL, aes(x = RecordEnd)) + 
  geom_histogram(bins = 13, fill = "blue", col = "white") + 
  theme_minimal()
# VehAge
ggplot(freMPL, aes(x = VehAge)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# Gender
ggplot(freMPL, aes(x = Gender)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# Maristat
ggplot(freMPL, aes(x = MariStat)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# SocioCateg
ggplot(freMPL, aes(x = SocioCateg)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
# VehUsage
ggplot(freMPL, aes(x = VehUsage)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# DrivAge
ggplot(freMPL, aes(x = DrivAge)) + 
  geom_histogram(bins = 30, fill = "blue", col = "white") + 
  theme_minimal()
# HasKmLimit (A numeric, 1 if there is a km limit for the policy, 0 otherwise)
ggplot(freMPL, aes(x = HasKmLimit)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# BonusMalus
ggplot(freMPL, aes(x = BonusMalus)) + 
  geom_histogram(bins = 30, fill = "blue", col = "white") + 
  theme_minimal()
# VehBody
ggplot(freMPL, aes(x = VehBody)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
# VehPrice (vehicle price from "A" (cheapest) to "Z" (most expensive))
ggplot(freMPL, aes(x = VehPrice)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# VehEngine
ggplot(freMPL, aes(x = VehEngine)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# VehEnergi
ggplot(freMPL, aes(x = VehEnergy)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# VehMaxSpeed
ggplot(freMPL, aes(x = VehMaxSpeed)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
# VehClass
ggplot(freMPL, aes(x = VehClass)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# RiskVar
ggplot(freMPL, aes(x = RiskVar)) + 
  geom_histogram(bins = 20, fill = "blue", col = "white") + 
  theme_minimal()
# Garage
ggplot(freMPL, aes(x = Garage)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# ClaimInd (Claim indicator of the guarantee. (this is not the claim number))
ggplot(freMPL, aes(x = ClaimInd)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
```

```{r}
# This below takes to long time: 
# subset_freMPL <- subset(freMPL, select = -c(SocioCateg, VehPrice)) # Too high cardinality
# sub_task = as_task_regr(subset_freMPL, target = "ClaimAmount")
# autoplot(sub_task, type = "pairs")
```

Responce variable: ClaimAmount
This variable looks heavytailed. 
```{r}
ggplot(freMPL, aes(x = ClaimAmount)) + 
  geom_histogram(bins = 20, fill = "blue", col = "white") + 
  theme_minimal()
```


# Noter: Hvilke tanker tænker vi at gennemgå?
Ridge regression, Lasso regression 
Spearman plot for at undersøge korrelationer mellem features

# Forsøg på at gentage det, vi lavede i øvelserne i Week2_Sol:

Define a task: 
```{r}
task = as_task_regr(freMPL, target = "ClaimAmount") # Regression task 
```

Put into train/test set
```{r}
set.seed(2026)
splits = partition(task, ratio = 0.8)

task_train = task$clone()$filter(splits$train)
task_test = task$clone()$filter(splits$test)
```

Define learner
```{r}
# my_learner = lrn("regr.rpart") # regression tree
# lrn("regr.rpart")$help() # Help
# Note: my_learner doesn't work because data contains characters, and rpart
# doesn't know how to handle this

task_train$feature_types

graph =
  po("colapply", applicator = as.factor, affect_columns = selector_type("character")) %>>%
  po("datefeatures") %>>%   # converts Date → numeric features
  po("encode") %>>%
  lrn("regr.rpart")

learner = as_learner(graph)

learner$train(task_train)
learner$predict(task_test)$score(msrs(c("regr.mse","regr.mae")))
```
The above i basically a manual holdout. Holdout below: 
```{r}
set.seed(2026)

resampling = rsmp("holdout", ratio = 0.8)
rr = resample(task, learner, resampling)

rr$aggregate(msrs(c("regr.mse","regr.mae")))
```
Notice how this gives the same result as the manual approch above.

Take a look at the model
```{r}
learner$model

learner$model$regr.rpart$model
```
Comments on the regression tree: 
* First split is on ClaimInt: Makes sence to split between if a claim i covered (is this what ClaimInt tells? And does it makes sence to use this for predicting? We don't know for a future claim if it is rejected or not. But if it means wheather a claim is covered by the insurance, then it might make sence to keep it in the model)
* Does it logically makes sence to split by the day (seasons with split by month might make better sence)
* Does it make sence to split by day of week? 
* Split by BonusMalus around 100 makes sence (under 100 good, over bad)
* Split by LicAge makes sence: But it predicts lower for "young" customers?? (the once 
who havent had a licence for a long time) (but split by 571.9 ~ 47,7 years of divers licence 
--> makes driver older than 47,7+18 = 65,7)


Featureless learner: 
For comparison we train a featureless model (no covariates). 
This model just predicts the mean (no gain of infomation from covariates). 
```{r}
baseline <- lrn("regr.featureless")$train(task_train)
baseline$predict(task_test)$score(msrs(c("regr.mse","regr.mae")))
```
The regression tree performs better than featureless.
Note that data is pre-processed: 
* colapply fixes characters
* datefeatures fixes dates
* encode turns factors into numbers

Why use pipelines instead of cleaning data yourself: 
* The pre-processing is learned on the train-data
* The same pre-processing is used on the test-data (no "knowledge" from test-data
is leaked to the pre-processing part)
* Det skaber en fejl i resultat, hvis testdata benyttes til at konvertere data. 

Why use rsmp() and resample(): 
* Shorter code - the split is handled by the functions, and thus no need to split data.
* The result is the same though




```{r}
MPL_data = freMPL

```

```{r}
summary(MPL_data)
```

```{r}
data_exp <- subset(MPL_data, Exposure == 1)
```

```{r}
unique(data_exp$RecordEnd)
```

```{r}
hist(data_exp$ClaimAmount, main ="Claim amount distribution",
    xlab ="Claim Amount",
    col ="blue")
```

```{r}
data_exp <- subset(data_exp, select = -c(RecordEnd, RecordBeg, ClaimInd))
```

```{r}
# Creating a new data set where few observations are lumped together as "other" to avoid problems with test-observations that are not not seen in the training data
data_exp_lump <- data_exp

factor_cols <- names(data_exp_lump)[sapply(data_exp_lump, function(x) is.factor(x) || is.character(x))]

for (col in factor_cols) {
  data_exp_lump[[col]] <- as.factor(data_exp_lump[[col]])
  data_exp_lump[[col]] <- fct_lump_min(data_exp_lump[[col]], min = 10)
  data_exp_lump[[col]] <- droplevels(data_exp_lump[[col]])
}
```

```{r}
unique(data_exp_lump$SocioCateg)
```


#################################
####### Testing Exercise solutions from week 2 on the data

# Define task

```{r}
task = as_task_regr(data_exp_lump, target = "ClaimAmount")
```

# Create train and test set

```{r}
set.seed(2026)
splits = partition(task, ratio = 0.8)
task_train = task$clone()$filter(splits$train)
task_test = task$clone()$filter(splits$test)
```

#### define learner: linear regression

```{r}
my_lr_learner = lrn("regr.lm")
```


##### Creates a pipeline that converts categorical features to numeric with one hot encoding, standardizes numeric features, and applies linear regression.

```{r}
graph_learner_lr = as_learner(
  po("encode", method = "one-hot") %>>%
  po("scale") %>>%
  my_lr_learner
)
```


######## result on test set

```{r}
graph_learner_lr$train(task_train)
graph_learner_lr$predict(task_test)$score(msr("regr.rmse"))
```

### baseline with no features

```{r}
baseline <- lrn("regr.featureless")$train(task_train)
baseline$predict(task_test)$score(msr("regr.rmse"))
```

#################################
### Cross Validation

#### define learner

### elastic net
```{r}
my_elasticnet_learner = lrn(
  "regr.glmnet", 
  standardize = FALSE,
  alpha = to_tune(0, 1),
  lambda = to_tune(p_dbl(log(1e-5), log(1e1), trafo = exp))
  # Alternatively lambda = to_tune(1e-5, 1e1, logscale=TRUE)
  )
```

#### create a graph that first one hot-encodes factors, and afterwards applies elastic net
```{r}
graph_learner_elastic_net = as_learner(
  po("encode", method = "one-hot") %>>%
    po("scale") %>>%
    my_elasticnet_learner
)
```

```{r}
set.seed(2026)
instance = tune(
  tuner = tnr("random_search", batch_size=20), ### tuning method
  #tuner = tnr("mbo"), ### tuning method
  task = task_train,
  learner = graph_learner_elastic_net,
  resampling = rsmp("cv", folds = 5), #### resampling method: 5-fold cross validation
  measures = msr("regr.rmse"), #### root mean squared error
  terminator = trm("evals", n_evals = 50) #### terminator
)
```

```{r}
as.data.table(instance$archive$data)
instance$result_learner_param_vals
```

```{r}
instance$result_y
```

# keep an untuned copy of the same graph for the auto_tuner example below
```{r}
graph_learner_elastic_net_auto = graph_learner_elastic_net$clone(deep = TRUE)
```

######## result on test set

```{r}
graph_learner_elastic_net$param_set$values = instance$result_learner_param_vals
graph_learner_elastic_net$train(task_train)
graph_learner_elastic_net$predict(task_test)$score(msr("regr.rmse"))

archive_dt = as.data.table(instance$archive)
```


### sparsity / accuracy tradeoff for a few alpha values
```{r}
alpha_values = c(0, 0.25, 0.5, 0.75, 1)
lambda_grid = exp(seq(log(1e-5), log(1e1), length.out = 15))
tradeoff = expand.grid(alpha = alpha_values, lambda = lambda_grid)
tradeoff$n_features = NA_real_
tradeoff$cv_error = NA_real_
```


```{r}
for (i in seq_len(nrow(tradeoff))) {
  learner_i = as_learner(
    po("encode", method = "one-hot") %>>% po("scale") %>>%
      lrn("regr.glmnet",
          standardize = FALSE,
          alpha = tradeoff$alpha[i],
          lambda = tradeoff$lambda[i])
  )

  set.seed(2026)
  rr_i = resample(task_train, learner_i$clone(deep = TRUE), rsmp("cv", folds = 5))
  tradeoff$cv_error[i] = rr_i$aggregate(msr("regr.rmse"))

  learner_i$train(task_train)
  beta = learner_i$model$regr.glmnet$model$beta
  tradeoff$n_features[i] = sum(as.matrix(beta) != 0)
}
```

# Plot results using ggplot2
```{r}
ggplot(data=tradeoff) +
  geom_point(aes(x=n_features, y=cv_error, color=as.factor(alpha)))
```

# Plot results using base-r
```{r}
score_range = range(tradeoff$cv_error)
point_cols = c("#1a9850", "#91cf60", "#ffffbf", "#fc8d59", "#d73027")
line_cols = c("#1b9e77", "#66a61e", "#d95f02", "#e7298a", "#7570b3")

score_breaks = seq(score_range[1], score_range[2], length.out = length(point_cols) + 1)
color_id = cut(tradeoff$cv_error, breaks = score_breaks, include.lowest = TRUE, labels = FALSE)
tradeoff$point_col = point_cols[color_id]
legend_labels = paste0(
  sprintf("%.3f", score_breaks[-length(score_breaks)]),
  " - ",
  sprintf("%.3f", score_breaks[-1])
)


plot(NA,
     xlim = range(lambda_grid),
     ylim = range(tradeoff$n_features),
     log = "x",
     xlab = "lambda",
     ylab = "Number of non-zero coefficients",
     main = "Sparsity / accuracy tradeoff")

for (j in seq_along(alpha_values)) {
  alpha_j = alpha_values[j]
  tradeoff_j = tradeoff[tradeoff$alpha == alpha_j, ]

  lines(tradeoff_j$lambda, tradeoff_j$n_features, col = line_cols[j], lwd = 2)
  points(tradeoff_j$lambda, tradeoff_j$n_features,
         pch = 21,
         bg = tradeoff_j$point_col,
         col = line_cols[j],
         cex = 1.1)
}

legend("topright",
       legend = paste("alpha =", alpha_values),
       col = line_cols,
       lwd = 2,
       pch = 21,
       pt.bg = "white",
       bty = "n")

legend("bottomleft",
       legend = legend_labels,
       fill = point_cols,
       title = "CV error",
       bty = "n",
       cex = 0.8)
```


### alternative: use auto_tuner already here
```{r}
at_test = auto_tuner(
  tuner = tnr("random_search", batch_size = 20),
  learner = graph_learner_elastic_net_auto,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 50)
)

set.seed(2026)
at_test$train(task_train)
at_test$predict(task_test)$score(msr("regr.rmse"))
at_test$tuning_instance$result_learner_param_vals
```


### see beta
```{r}
graph_learner_elastic_net$model$regr.glmnet$model$beta
```



#################################
### Nested Cross Validation

```{r}
at = auto_tuner(
  tuner = tnr("random_search", batch_size=20),
  learner = graph_learner_elastic_net,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 50)
)

set.seed(2026)
rr = resample(task, at , rsmp("cv", folds = 5))
rr
rr$score(msr("regr.rmse"))[, .(iteration, regr.rmse)]
rr$aggregate(msr("regr.rmse"))

```

######## result on test set

```{r}
graph_learner_elastic_net$param_set$values = instance$result_learner_param_vals
graph_learner_elastic_net$train(task_train)
graph_learner_elastic_net$predict(task_test)$score(msr("regr.rmse"))

archive_dt = as.data.table(instance$archive)
```

#kunde 
1 60 - - 0 Male Alone CSP50 Priave + office 30 0 150 sport utility V Direct injection overpowered disel 1900-200 H risk var 20 collective garage








