---
title: "Aflevering"
output: pdf_document
date: "2026-05-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Library
```{r}
library(mlr3)
library(mlr3learners)
library(mlr3tuning)
library(mlr3mbo)
library(mlr3pipelines)
library(paradox)
library(readr)
library(tidyverse)
library(mlr3viz)
library(mlr3torch)
library(torch)
library(future)
library(forcats)
library(mlr3extralearners)
library(OpenML)
library(mgcv)
library(kknn)
library(lightgbm)
library(Metrics)
library(xgboost)
library(iml)
```

# Load data

```{r}
freMPL <- read_csv("freMPL.csv")
```

# Pre-processing

We remove ClaimInd from the dataset, since we want to predict a price for a given customer, and we don't know, if the customer will have a claim.
Remove ...1 since it is just row_numbers.
```{r}
MPL_data <- subset(freMPL, select = -c(ClaimInd,...1))
```

We replace null-values from RecordEnd, so we don't have to deal with null values.
```{r}
MPL_data$RecordEnd[is.na(MPL_data$RecordEnd)] <- as.Date("2005-01-01")
```

We replace all date-values with numeric values, because many models cannot handle date-values.
```{r}
MPL_data$RecordBeg <- as.numeric(MPL_data$RecordBeg)
MPL_data$RecordEnd <- as.numeric(MPL_data$RecordEnd)
```

Ordered factors and factors.
```{r}
# VehAge is an ordered variable
MPL_data$VehAge <- factor(MPL_data$VehAge, levels = c("0", "1", "2", "3", "4", "5", "6-7", "8-9", "10+"))

# VehMaxSpeed is an ordered variable
MPL_data$VehMaxSpeed <- factor(MPL_data$VehMaxSpeed,
                             levels = c("1-130 km/h", "130-140 km/h", "140-150 km/h",
                                        "150-160 km/h", "160-170 km/h", "170-180 km/h", 
                                        "180-190 km/h", "190-200 km/h", "200-220 km/h",
                                        "220+ km/h"))

# HasKmLimit
MPL_data$HasKmLimit <- factor(MPL_data$HasKmLimit)
```

Creating a new data set where few observations are lumped together as "other" to avoid problems with test-observations that are not not seen in the training data. 
```{r}
MPL_data_lump <- MPL_data

factor_cols <- names(MPL_data_lump)[sapply(MPL_data_lump, function(x) is.factor(x) || is.character(x))]

for (col in factor_cols) {
  MPL_data_lump[[col]] <- as.factor(MPL_data_lump[[col]])
  MPL_data_lump[[col]] <- fct_lump_min(MPL_data_lump[[col]], min = 10)
  MPL_data_lump[[col]] <- droplevels(MPL_data_lump[[col]])
}

categorical_vars <- c("VehAge", "Gender", "MariStat", "SocioCateg", "VehUsage", "VehBody", "VehPrice", "VehEngine", "VehEnergy", "VehMaxSpeed", "VehClass", "Garage") # VehMax Speed should perhaps be converted to a numeric variable by taking eg the max number in the observations given interval

MPL_data_lump[categorical_vars] <- lapply(MPL_data_lump[categorical_vars], factor)
```

Final data: 
```{r}
summary(MPL_data_lump)
```

# Generisk kunde 1: 
Gennemsnits kunde
```{r}
gns_kunde <- MPL_data_lump[1,]

gns_kunde$Exposure <- 1  
gns_kunde$LicAge <- as.numeric(as.integer(mean(MPL_data_lump$LicAge)))
gns_kunde$RecordBeg <- as.numeric(as.Date("2004-01-01"))
gns_kunde$RecordEnd <- as.numeric(as.Date("2005-01-01"))
gns_kunde$VehAge <- factor(names(sort(table(MPL_data_lump$VehAge), decreasing = TRUE))[1],
                           levels = levels(MPL_data_lump$VehAge))
gns_kunde$Gender <- factor(names(sort(table(MPL_data_lump$Gender), decreasing = TRUE))[1],
                           levels = levels(MPL_data_lump$Gender))
gns_kunde$MariStat <- factor(names(sort(table(MPL_data_lump$MariStat), decreasing = TRUE))[1],
                             levels = levels(MPL_data_lump$MariStat))
gns_kunde$SocioCateg <- factor(names(sort(table(MPL_data_lump$SocioCateg), decreasing = TRUE))[1],
                               levels = levels(MPL_data_lump$SocioCateg))
gns_kunde$VehUsage <- factor(names(sort(table(MPL_data_lump$VehUsage), decreasing = TRUE))[1],
                             levels = levels(MPL_data_lump$VehUsage))
gns_kunde$DrivAge <- as.numeric(as.integer(mean(MPL_data_lump$DrivAge)))
gns_kunde$HasKmLimit <- as.numeric(names(sort(table(MPL_data_lump$HasKmLimit), decreasing = TRUE))[1])
gns_kunde$BonusMalus <- as.numeric(as.integer(mean(MPL_data_lump$BonusMalus)))
gns_kunde$VehBody <- factor(names(sort(table(MPL_data_lump$VehBody), decreasing = TRUE))[1],
                            levels = levels(MPL_data_lump$VehBody))
gns_kunde$VehPrice <- factor(names(sort(table(MPL_data_lump$VehPrice), decreasing = TRUE))[1],
                             levels = levels(MPL_data_lump$VehPrice))
gns_kunde$VehEngine <- factor(names(sort(table(MPL_data_lump$VehEngine), decreasing = TRUE))[1],
                              levels = levels(MPL_data_lump$VehEngine))
gns_kunde$VehEnergy <- factor(names(sort(table(MPL_data_lump$VehEnergy), decreasing = TRUE))[1],
                              levels = levels(MPL_data_lump$VehEnergy))
gns_kunde$VehMaxSpeed <- factor(names(sort(table(MPL_data_lump$VehMaxSpeed), decreasing = TRUE))[1],
                                levels = levels(MPL_data_lump$VehMaxSpeed))
gns_kunde$VehClass <- factor(names(sort(table(MPL_data_lump$VehClass), decreasing = TRUE))[1],
                             levels = levels(MPL_data_lump$VehClass))
gns_kunde$ClaimAmount <- mean(MPL_data_lump$ClaimAmount)
gns_kunde$RiskVar <- as.numeric(as.integer(mean(MPL_data_lump$RiskVar)))
gns_kunde$Garage <- factor(names(sort(table(MPL_data_lump$Garage), decreasing = TRUE))[1],
                           levels = levels(MPL_data_lump$Garage))
```

# Generisk kunde 2: 
Ekstrem kunde

#### TASK and SPLIT 

# Task
```{r}
task = as_task_regr(MPL_data_lump, target = "ClaimAmount")
```

```{r}
set.seed(2026)

split <- partition(task, ratio = 0.8)

train_set <- split$train
test_set <- split$test
```


#### BASELINE MODEL FOR COMPARISON ######

```{r}
baseline <- lrn("regr.featureless")
```

```{r}
baseline$train(task, row_ids = split$train)
```

```{r}
pred_baseline <- baseline$predict(task, row_ids = split$test)
```

```{r}
pred_baseline$score(msr("regr.rmse"))
pred_baseline$score(msr("regr.mae"))
pred_baseline$score(msr("regr.rsq"))
```

##### Tree

```{r}
categorical_selector = selector_type(c("factor", "ordered"))

make_encoded_learner = function(learner) {
  as_learner(
    po("colapply", applicator = as.factor, affect_columns = selector_type("character")) %>>%
      po("datefeatures") %>>%
      po("encode", method = "treatment",
         affect_columns = categorical_selector, id = "binary_enc") %>>%
      learner
  )
}
```

Defult tree
```{r}
lrn_tree = make_encoded_learner(lrn("regr.rpart"))
lrn_tree$train(task, row_ids = split$train)

plot(lrn_tree$model$regr.rpart$model, compress = TRUE, margin = 0.1)
text(lrn_tree$model$regr.rpart$model, use.n = TRUE, cex = 0.7)

lrn_tree$predict(task, row_ids = split$test)$score(msrs(c("regr.mse","regr.rmse")))
```

Define learner - full tree
```{r}
full_tree = make_encoded_learner(lrn("regr.rpart", cp = 0))
full_tree$train(task, row_ids = split$train)

plot(full_tree$model$regr.rpart$model, compress = TRUE, margin = 0.1)
text(full_tree$model$regr.rpart$model, use.n = TRUE, cex = 0.2)

full_tree$predict(task, row_ids = split$test)$score(msrs(c("regr.mse","regr.rmse")))
```

Optimal cp
```{r}
graph_learner_cart_cv = make_encoded_learner(lrn("regr.rpart", cp = 0, xval = 5))
graph_learner_cart_cv$train(task, row_ids = split$train)

cart_trained_cv <- graph_learner_cart_cv$model$regr.rpart$model

rpart::plotcp(cart_trained_cv)
rpart::printcp(cart_trained_cv)

cp_table = cart_trained_cv$cptable
min_row = which.min(cp_table[, "xerror"])
one_se_threshold = cp_table[min_row, "xerror"] + cp_table[min_row, "xstd"]

selected_row = which(cp_table[, "xerror"] <= one_se_threshold)[1]
selected_cp = cp_table[selected_row, "CP"]
selected_cp
```

```{r}
selected_cp
pruned_tree = make_encoded_learner(lrn("regr.rpart", cp = 0.02459138))
pruned_tree$train(task, row_ids = split$train)

plot(pruned_tree$model$regr.rpart$model, compress = TRUE, margin = 0.1)
text(pruned_tree$model$regr.rpart$model, use.n = TRUE, cex = 0.5)

pruned_tree$predict(task, row_ids = split$test)$score(msrs(c("regr.mse","regr.rmse")))
```


##### DEFINING KNN MODEL #####

```{r}
graph_knn <- 
  po("encode", method = "one-hot") %>>%
  po("scale") %>>%
  lrn("regr.kknn")
```

```{r}
learner_knn <- as_learner(graph_knn)
```

```{r}
learner_knn$train(task, row_ids = train_set)
```

```{r}
prediction_knn <- learner_knn$predict(task, row_ids = test_set)
```

```{r}
prediction_knn$score(msr("regr.rmse"))
prediction_knn$score(msr("regr.mae"))
prediction_knn$score(msr("regr.rsq"))
```

#### KNN TUNING ####

```{r}
learner_knn <- lrn(
  "regr.kknn",
  k = to_tune(1, 50)
)
```

```{r}
graph_knn <- 
  po("encode", method = "one-hot") %>>%
  po("scale") %>>%
  learner_knn

graph_learner_knn <- as_learner(graph_knn)
```

```{r}
at_knn <- auto_tuner(
  tuner = tnr("random_search"),
  learner = graph_learner_knn,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 20)
)
```

```{r}
at_knn$train(task, row_ids = train_set)
```

```{r}
at_knn$tuning_result
```

```{r}
pred_knn <- at_knn$predict(task, row_ids = test_set)

pred_knn$score(msr("regr.rmse"))
pred_knn$score(msr("regr.mae"))
pred_knn$score(msr("regr.rsq"))
```

##### DEFAULT RF MODEL #####

```{r}
task_rf <- TaskRegr$new(
  id = "claim_rf",
  backend = MPL_data_lump,
  target = "ClaimAmount"
)
```

```{r}
learner_rf <- lrn(
  "regr.ranger",             #ranger implements Random Forest
  importance = "impurity",
  predict_type = "response"
)
```

```{r}
learner_rf$train(task, row_ids = train_set)
```

```{r}
pred_rf <- learner_rf$predict(task, row_ids = test_set)

pred_rf$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

#### TARGET ENCODING ######

```{r}
target_encoder <- po(
  "encodeimpact",
  affect_columns = selector_type(c("factor", "character"))
)
```

```{r}
learner_rf <- lrn(
  "regr.ranger",
  importance = "impurity",
  predict_type = "response"
)
```

```{r}
graph_rf_te <- target_encoder %>>% learner_rf

graph_learner_rf_te <- as_learner(graph_rf_te)
```

```{r}
graph_learner_rf_te$train(
  task,
  row_ids = train_set
)
```

```{r}
pred_rf_te <- graph_learner_rf_te$predict(
  task,
  row_ids = test_set
)
```

```{r}
pred_rf_te$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

##### TUNING ######

```{r}
learner_rf_tuned <- as_learner(
  po("encode", method = "one-hot") %>>%
    lrn("regr.ranger",
                importance = "impurity",
                predict_type = "response",
                num.trees = to_tune(c(500:1500)),
                min.node.size = to_tune(c(1:20)),
                mtry = to_tune(c(2:12)),
                sample.fraction = to_tune(p_dbl(lower = 0.5, upper = 1))
                )
)
```

```{r}
at_rf <- auto_tuner(
  tuner = tnr("random_search"),
  learner = learner_rf_tuned,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 25)
)
```

```{r}
at_rf$train(task, row_ids = train_set)
```

```{r}
at_rf$tuning_result
```

```{r}
pred_tuned <- at_rf$predict(task, row_ids = test_set)

pred_tuned$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

#### TUNING WITH NEW NODE SIZES ####

```{r}
learner_rf_tuned_2 <- as_learner(
  po("encode", method = "one-hot") %>>%
    lrn("regr.ranger",
                importance = "impurity",
                predict_type = "response",
                num.trees = to_tune(c(500:1500)),
                min.node.size = to_tune(c(100:2000)),
                mtry = to_tune(c(2:12)),
                sample.fraction = to_tune(p_dbl(lower = 0.5, upper = 1))
                )
)
```

```{r}
at_rf_2 <- auto_tuner(
  tuner = tnr("random_search"),
  learner = learner_rf_tuned_2,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 25)
)
```

```{r}
at_rf_2$train(task, row_ids = train_set)
```

```{r}
at_rf_2$tuning_result
```

```{r}
pred_tuned_2 <- at_rf_2$predict(task, row_ids = test_set)

pred_tuned_2$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

### RESULTS RF ###

```{r}
results_rf <- data.frame(
  Model = c("Baseline", "RF Default", "RF Target Encoded", "RF Tuned", "RF Tuned Node"),
  RMSE = c(pred_baseline$score(msr("regr.rmse")), pred_rf$score(msr("regr.rmse")), pred_rf_te$score(msr("regr.rmse")), pred_tuned$score(msr("regr.rmse")), pred_tuned_2$score(msr("regr.rmse"))),
  MAE = c(pred_baseline$score(msr("regr.mae")), pred_rf$score(msr("regr.mae")), pred_rf_te$score(msr("regr.mae")), pred_tuned$score(msr("regr.mae")), pred_tuned_2$score(msr("regr.mae"))),
  RSQ = c(pred_baseline$score(msr("regr.rsq")), pred_rf$score(msr("regr.rsq")), pred_rf_te$score(msr("regr.rsq")), pred_tuned$score(msr("regr.rsq")), pred_tuned_2$score(msr("regr.rsq")))
)
results_rf
```

##### DEFAULT XGBoost MODEL #####

```{r}
learner_xgb_default <- as_learner(
  po("encode", method = "one-hot") %>>%
    lrn(
      "regr.xgboost",
      objective = "reg:squarederror",
      eval_metric = "rmse"
    )
)
```

```{r}
learner_xgb_default$train(
  task,
  row_ids = train_set
)
```

```{r}
pred_xgb_default <- learner_xgb_default$predict(
  task,
  row_ids = test_set
)
```

```{r}
pred_xgb_default$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

#### TARGET ENCODING #####

```{r}
learner_xgb_te <- as_learner(
  po("encodeimpact") %>>% #encodeimpact performs target encoding
    lrn(
      "regr.xgboost",
      objective = "reg:squarederror",
      eval_metric = "rmse"
    )
)
```

```{r}
learner_xgb_te$train(
  task,
  row_ids = train_set
)
```

```{r}
pred_xgb_te <- learner_xgb_te$predict(
  task,
  row_ids = test_set
)
```

```{r}
pred_xgb_te$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

The performance of the two models with the different encoding methods was very similar. However, the one-hot encoding lead to a slightly better model performance and therefore, we proceed with tuning that models hyperparameters.

##### TUNING #####

```{r}
learner_xgb_tuned <- as_learner(
  po("encode", method = "one-hot") %>>%
    lrn(
      "regr.xgboost",
      eta = to_tune(p_dbl(0.01, 0.3)),
      max_depth = to_tune(c(1:10)),
      nrounds = to_tune(c(100:1000))
    )
)
```

```{r}
at_xgb <- auto_tuner(
  tuner = tnr("random_search"),
  learner = learner_xgb_tuned,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 50)
)
```

```{r}
at_xgb$train(task, row_ids = train_set)
```
 
```{r}
pred_xgb_tuned <- at_xgb$predict(task, row_ids = test_set)

pred_xgb_tuned$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

###### RESULTS XGBoost ######

```{r}
results_xgb <- data.frame(
  Model = c("Baseline", "XGB Default", "XGB Target Encoded", "XGB Tuned"),
  RMSE = c(pred_baseline$score(msr("regr.rmse")), pred_xgb_default$score(msr("regr.rmse")), pred_xgb_te$score(msr("regr.rmse")), pred_xgb_tuned$score(msr("regr.rmse"))),
  MAE = c(pred_baseline$score(msr("regr.mae")), pred_xgb_default$score(msr("regr.mae")), pred_xgb_te$score(msr("regr.mae")), pred_xgb_tuned$score(msr("regr.mae"))),
  RSQ = c(pred_baseline$score(msr("regr.rsq")), pred_xgb_default$score(msr("regr.rsq")), pred_xgb_te$score(msr("regr.rsq")), pred_xgb_tuned$score(msr("regr.rsq")))
)
results_xgb
```

##### DEFAULT LIGHTGBM MODEL #####

```{r}
learner_lgb_default <- as_learner(
  po("encode", method = "one-hot") %>>%
    lrn("regr.lightgbm",
        objective = "regression",
        verbose = -1
    )
)
```

```{r}
learner_lgb_default$train(
  task,
  row_ids = train_set
)
```

```{r}
pred_lgb_default <- learner_lgb_default$predict(
  task,
  row_ids = test_set
)
```

```{r}
pred_lgb_default$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

#### TARGET ENCODING ####

```{r}
learner_lgb_te <- as_learner(
  po("encodeimpact") %>>%
    lrn(
      "regr.lightgbm",
      objective = "regression",
      verbose = -1
    )
)
```

```{r}
learner_lgb_te$train(
  task,
  row_ids = train_set
)
```

```{r}
pred_lgb_te <- learner_lgb_te$predict(
  task,
  row_ids = test_set
)
```

```{r}
pred_lgb_te$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

Once again, the target encoding does not improve performance, so we proceed with tuning the hyperparameters of the default model.

##### TUNING ####

```{r}
learner_lgb_tuned <- as_learner(
  po("encode", method = "one-hot") %>>%
    lrn("regr.lightgbm",
        learning_rate = to_tune(p_dbl(0.01, 0.1)),
        num_leaves = to_tune(c(20:80)),
        max_depth = to_tune(c(1:10)),
        min_data_in_leaf = to_tune(c(20:80)),
        num_iterations = to_tune(c(100:500))
    )
)
```

```{r}
at_lgb <- auto_tuner(
  tuner = tnr("random_search"),
  learner = learner_lgb_tuned,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 50)
)
```

```{r}
at_lgb$train(task, row_ids = train_set)
```

```{r}
pred_lgb_tuned <- at_lgb$predict(task, row_ids = test_set)

pred_lgb_tuned$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

#### TUNING WITH LARGER LEAVES ###

```{r}
learner_lgb_tuned_2 <- as_learner(
  po("encode", method = "one-hot") %>>%
    lrn("regr.lightgbm",
        learning_rate = to_tune(p_dbl(0.01, 0.1)),
        num_leaves = to_tune(c(20:80)),
        max_depth = to_tune(c(1:10)),
        min_data_in_leaf = to_tune(c(100:2000)),
        num_iterations = to_tune(c(100:500))
    )
)
```

```{r}
at_lgb_2 <- auto_tuner(
  tuner = tnr("random_search"),
  learner = learner_lgb_tuned,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 50)
)
```

```{r}
at_lgb_2$train(task, row_ids = train_set)
```

```{r}
pred_lgb_tuned_2 <- at_lgb_2$predict(task, row_ids = test_set)

pred_lgb_tuned_2$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

###### RESULTS LGBM ######

```{r}
results_lgb <- data.frame(
  Model = c("Baseline", "LGBM Default", "LGBM Target Encoded", "LGBM Tuned", "LGBM Tuned Leaf size"),
  RMSE = c(pred_baseline$score(msr("regr.rmse")), pred_lgb_default$score(msr("regr.rmse")), pred_lgb_te$score(msr("regr.rmse")), pred_lgb_tuned$score(msr("regr.rmse")), pred_lgb_tuned_2$score(msr("regr.rmse"))),
  MAE = c(pred_baseline$score(msr("regr.mae")), pred_lgb_default$score(msr("regr.mae")), pred_lgb_te$score(msr("regr.mae")), pred_lgb_tuned$score(msr("regr.mae")), pred_lgb_tuned_2$score(msr("regr.mae"))),
  RSQ = c(pred_baseline$score(msr("regr.rsq")), pred_lgb_default$score(msr("regr.rsq")), pred_lgb_te$score(msr("regr.rsq")), pred_lgb_tuned$score(msr("regr.rsq")), pred_lgb_tuned_2$score(msr("regr.rsq")))
)
results_lgb
```

##### DEFAULT NEURAL NETWORK MODEL #####

```{r}
# MLP (Multi layer perceptron): Adam optimizer. SGD on mini batches (batch size = 32). See mlr_learners$get("regr.mlp")
graph_nn =
  po("encode", method = "one-hot") %>>%
  po("scale") %>>%
  lrn(
    "regr.mlp",
    epochs = 100,
    batch_size = 32,
    patience = 20, # Stopping process after 10 epochs without any improvement on validation set.
    measures_valid = msr("regr.rmse")
  )
```

```{r}
learner_nn = as_learner(graph_nn)
```

```{r}
set_validate(learner_nn, validate = 0.2) 
# Splitting the training data into new training and validation sets. This is used for early stopping.
```

```{r}
learner_nn$train(
  task,
  row_ids = train_set
)
```

```{r}
pred_nn = learner_nn$predict(
  task,
  row_ids = test_set
)
```

```{r}
pred_nn$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

###### RESULTS NN ######

```{r}
results_nn <- data.frame(
  Model = c("Baseline", "NN Default"),
  RMSE = c(pred_baseline$score(msr("regr.rmse")), pred_nn$score(msr("regr.rmse"))),
  MAE = c(pred_baseline$score(msr("regr.mae")), pred_nn$score(msr("regr.mae"))),
  RSQ = c(pred_baseline$score(msr("regr.rsq")), pred_nn$score(msr("regr.rsq")))
)
results_nn
```

We start by making a matrix for the results we want 
```{r}
results <- matrix(0, nrow=5, ncol=3)
rownames(results) = c("LicAge", "Exposure","DrivAge","BonusMalus","VehPrice")
colnames(results) = c("regr.rmse", "regr.mae", "regr.rsq")
```

Now we make a model for each of our five chosen coefficient, where we focus on one off the coefficients and put all of the other together.

##LicAge
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(MPL_data_lump), c("LicAge", "ClaimAmount")))
task_LicAge_on_allother = as_task_regr(MPL_data_lump, target="LicAge")$select(setdiff(names(MPL_data_lump), c("LicAge", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_LicAge_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=LicAge ~ Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

cross_fit = function(task, learner, K){
  set.seed(2026)
  outer_resampling = rsmp("cv", folds=K)
  rr = resample(task, learner, outer_resampling)
  pred = rr$prediction() 
  pred$truth - pred$response
}

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
LicAge_residuals_smooth = cross_fit(task_LicAge_on_allother, lrn_LicAge_on_allother_smooth, K=2)

lm_amount_on_LicAge_smooth = lm(amount_residuals_smooth ~ LicAge_residuals_smooth)

results[1,] <- c(sqrt(mean(lm_amount_on_LicAge_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_LicAge_smooth)),
summary(lm_amount_on_LicAge_smooth)$r.squared)
```
##Exposure 
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(MPL_data_lump), c("Exposure", "ClaimAmount")))
task_Exposure_on_allother = as_task_regr(MPL_data_lump, target="Exposure")$select(setdiff(names(MPL_data_lump), c("Exposure", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_Exposure_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=Exposure ~ LicAge + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
Exposure_residuals_smooth = cross_fit(task_Exposure_on_allother, lrn_Exposure_on_allother_smooth, K=2)

lm_amount_on_Exposure_smooth = lm(amount_residuals_smooth ~ Exposure_residuals_smooth)

results[2,] <- c(sqrt(mean(lm_amount_on_Exposure_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_Exposure_smooth)),
summary(lm_amount_on_Exposure_smooth)$r.squared)
```
##DrivAge
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(MPL_data_lump), c("DrivAge", "ClaimAmount")))
task_DrivAge_on_allother = as_task_regr(MPL_data_lump, target="DrivAge")$select(setdiff(names(MPL_data_lump), c("DrivAge", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_DrivAge_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=DrivAge ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
DrivAge_residuals_smooth = cross_fit(task_DrivAge_on_allother, lrn_DrivAge_on_allother_smooth, K=2)

lm_amount_on_DrivAge_smooth = lm(amount_residuals_smooth ~ DrivAge_residuals_smooth)

results[3,] <- c(sqrt(mean(lm_amount_on_DrivAge_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_DrivAge_smooth)),
summary(lm_amount_on_DrivAge_smooth)$r.squared)
```
test
##BonusMalus
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(MPL_data_lump), c("BonusMalus", "ClaimAmount")))
task_BonusMalus_on_allother = as_task_regr(MPL_data_lump, target="BonusMalus")$select(setdiff(names(MPL_data_lump), c("BonusMalus", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_BonusMalus_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=BonusMalus ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
BonusMalus_residuals_smooth = cross_fit(task_BonusMalus_on_allother, lrn_BonusMalus_on_allother_smooth, K=2)

lm_amount_on_BonusMalus_smooth = lm(amount_residuals_smooth ~ BonusMalus_residuals_smooth)

results[4,] <- c(sqrt(mean(lm_amount_on_BonusMalus_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_BonusMalus_smooth)),
summary(lm_amount_on_BonusMalus_smooth)$r.squared)
```
##VehPrice 
```{r}
MPL_data_lump$VehPrice <- as.numeric(MPL_data_lump$VehPrice) #Can't have factor as a response variable

task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(MPL_data_lump), c("VehPrice", "ClaimAmount")))
task_VehPrice_on_allother = as_task_regr(MPL_data_lump, target="VehPrice")$select(setdiff(names(MPL_data_lump), c("VehPrice", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_VehPrice_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=VehPrice ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
VehPrice_residuals_smooth = cross_fit(task_VehPrice_on_allother, lrn_VehPrice_on_allother_smooth, K=2)

lm_amount_on_VehPrice_smooth = lm(amount_residuals_smooth ~ VehPrice_residuals_smooth)

results[5,] <- c(sqrt(mean(lm_amount_on_VehPrice_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_VehPrice_smooth)),
summary(lm_amount_on_VehPrice_smooth)$r.squared)

MPL_data_lump$VehPrice <- as.factor(MPL_data_lump$VehPrice) #Making it a factor again for later
```

## Shapley values 

```{r}
predict_fun <- function(model,task){ 
  Predictor$new(
    model = model,
    data = task$data(cols = task$feature_names),
    y = task$data(cols = task$target_names)[[1]]
    )
}

shapley_values <- function(model,task,new_data,n=1000){
  Shapley$new(
    predictor = predict_fun(model,task),
    x.interest = new_data,
    sample.size = n
  )$results$phi
}
```

Shapley values for gns_kunde
Random forrest second model made with tuning
```{r}
df <- 
  data.frame(
    x = task$feature_names,
    y = shapley_values(learner_rf_tuned_2,task,gns_kunde,1000)
  )


ggplot(df, aes(x=y,y=x)) +
  geom_col() +
  labs(x="Shapley Value", y = "Feature", title = "Gns_kunde")
```

## PDP and ALE

```{r}
explainer_fun <- function(learner_trained, task){
  explainer = DALEXtra::explain_mlr3(
    learner_trained,
    data = task$data(cols = task$feature_names),
    y = task$data(cols = task$target_names)[[1]],
    colorize = FALSE
  )
}
```

PDP and ALE for random forrest
```{r}
explainer <- explainer_fun(learner_rf_tuned_2, task)

pdps <- model_profile(explainer, type="partial")
ales <- model_profile(explainer, type="accumulated")

plot(pdps)
plot(ales)
```


