---
title: "Aflevering"
output: pdf_document
date: "2026-05-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Library
```{r}
library(mlr3)
library(mlr3learners)
library(mlr3tuning)
library(mlr3mbo)
library(mlr3pipelines)
library(paradox)
library(readr)
library(tidyverse)
library(mlr3viz)
library(mlr3torch)
library(torch)
library(future)
library(forcats)
library(mlr3extralearners)
library(OpenML)
library(mgcv)
library(kknn)
library(lightgbm)
library(Metrics)
```

# Load data

```{r}
freMPL <- read_csv("freMPL.csv")
```

# Pre-processing

We remove ClaimInd from the dataset, since we want to predict a price for a given customer, and we don't know, if the customer will have a claim.
Remove ...1 since it is just row_numbers.
```{r}
MPL_data <- subset(freMPL, select = -c(ClaimInd,...1))
```

We replace null-values from RecordEnd, so we don't have to deal with null values.
```{r}
MPL_data$RecordEnd[is.na(MPL_data$RecordEnd)] <- as.Date("2005-01-01")
```

We replace all date-values with numeric values, because many models cannot handle date-values.
```{r}
MPL_data$RecordBeg <- as.numeric(MPL_data$RecordBeg)
MPL_data$RecordEnd <- as.numeric(MPL_data$RecordEnd)
```

Ordered factors and factors.
```{r}
# VehAge is an ordered variable
freMPL$VehAge <- factor(freMPL$VehAge, levels = c("0", "1", "2", "3", "4", "5", "6-7", "8-9", "10+"))

# VehMaxSpeed is an ordered variable
freMPL$VehMaxSpeed <- factor(freMPL$VehMaxSpeed,
                             levels = c("1-130 km/h", "130-140 km/h", "140-150 km/h",
                                        "150-160 km/h", "160-170 km/h", "170-180 km/h", 
                                        "180-190 km/h", "190-200 km/h", "200-220 km/h",
                                        "220+ km/h"))

# HasKmLimit
freMPL$HasKmLimit <- factor(freMPL$HasKmLimit)
```

Creating a new data set where few observations are lumped together as "other" to avoid problems with test-observations that are not not seen in the training data. 
```{r}
MPL_data_lump <- MPL_data

factor_cols <- names(MPL_data_lump)[sapply(MPL_data_lump, function(x) is.factor(x) || is.character(x))]

for (col in factor_cols) {
  MPL_data_lump[[col]] <- as.factor(MPL_data_lump[[col]])
  MPL_data_lump[[col]] <- fct_lump_min(MPL_data_lump[[col]], min = 10)
  MPL_data_lump[[col]] <- droplevels(MPL_data_lump[[col]])
}

categorical_vars <- c("VehAge", "Gender", "MariStat", "SocioCateg", "VehUsage", "VehBody", "VehPrice", "VehEngine", "VehEnergy", "VehMaxSpeed", "VehClass", "Garage") # VehMax Speed should perhaps be converted to a numeric variable by taking eg the max number in the observations given interval

MPL_data_lump[categorical_vars] <- lapply(MPL_data_lump[categorical_vars], factor)
```

Final data: 
```{r}
summary(MPL_data_lump)
```

# Generisk kunde 1: 
Gennemsnits kunde
```{r}
gns_kunde <- MPL_data_lump[1,]

gns_kunde$Exposure <- 1  
gns_kunde$LicAge <- as.numeric(as.integer(mean(MPL_data_lump$LicAge)))
gns_kunde$RecordBeg <- as.Date("2004-01-01")
gns_kunde$RecordEnd <- as.Date("2005-01-01")
gns_kunde$VehAge <- factor(names(sort(table(MPL_data_lump$VehAge), decreasing = TRUE))[1],
                           levels = levels(MPL_data_lump$VehAge))
gns_kunde$Gender <- factor(names(sort(table(MPL_data_lump$Gender), decreasing = TRUE))[1],
                           levels = levels(MPL_data_lump$Gender))
gns_kunde$MariStat <- factor(names(sort(table(MPL_data_lump$MariStat), decreasing = TRUE))[1],
                             levels = levels(MPL_data_lump$MariStat))
gns_kunde$SocioCateg <- factor(names(sort(table(MPL_data_lump$SocioCateg), decreasing = TRUE))[1],
                               levels = levels(MPL_data_lump$SocioCateg))
gns_kunde$VehUsage <- factor(names(sort(table(MPL_data_lump$VehUsage), decreasing = TRUE))[1],
                             levels = levels(MPL_data_lump$VehUsage))
gns_kunde$DrivAge <- as.numeric(as.integer(mean(MPL_data_lump$DrivAge)))
gns_kunde$HasKmLimit <- as.numeric(names(sort(table(MPL_data_lump$HasKmLimit), decreasing = TRUE))[1])
gns_kunde$BonusMalus <- as.numeric(as.integer(mean(MPL_data_lump$BonusMalus)))
gns_kunde$VehBody <- factor(names(sort(table(MPL_data_lump$VehBody), decreasing = TRUE))[1],
                            levels = levels(MPL_data_lump$VehBody))
gns_kunde$VehPrice <- factor(names(sort(table(MPL_data_lump$VehPrice), decreasing = TRUE))[1],
                             levels = levels(MPL_data_lump$VehPrice))
gns_kunde$VehEngine <- factor(names(sort(table(MPL_data_lump$VehEngine), decreasing = TRUE))[1],
                              levels = levels(MPL_data_lump$VehEngine))
gns_kunde$VehEnergy <- factor(names(sort(table(MPL_data_lump$VehEnergy), decreasing = TRUE))[1],
                              levels = levels(MPL_data_lump$VehEnergy))
gns_kunde$VehMaxSpeed <- factor(names(sort(table(MPL_data_lump$VehMaxSpeed), decreasing = TRUE))[1],
                                levels = levels(MPL_data_lump$VehMaxSpeed))
gns_kunde$VehClass <- factor(names(sort(table(MPL_data_lump$VehClass), decreasing = TRUE))[1],
                             levels = levels(MPL_data_lump$VehClass))
gns_kunde$ClaimAmount <- mean(MPL_data_lump$ClaimAmount)
gns_kunde$RiskVar <- as.numeric(as.integer(mean(MPL_data_lump$RiskVar)))
gns_kunde$Garage <- factor(names(sort(table(MPL_data_lump$Garage), decreasing = TRUE))[1],
                           levels = levels(MPL_data_lump$Garage))
```

# Generisk kunde 2: 
Ekstrem kunde

#### TASK and SPLIT 

# Task
```{r}
task = as_task_regr(MPL_data_lump, target = "ClaimAmount")
```

```{r}
task <- TaskRegr$new(
  id = "claims",
  backend = MPL_data_lump,
  target = "ClaimAmount"
)
```

```{r}
set.seed(2026)

split <- partition(task, ratio = 0.8)

train_set <- split$train
test_set <- split$test
```


#### BASELINE MODEL FOR COMPARISON ######

```{r}
baseline <- lrn("regr.featureless")
```

```{r}
baseline$train(task, row_ids = split$train)
```

```{r}
pred_baseline <- baseline$predict(task, row_ids = split$test)
```

```{r}
pred_baseline$score(msr("regr.rmse"))
pred_baseline$score(msr("regr.mae"))
pred_baseline$score(msr("regr.rsq"))
```

##### DEFINING KNN MODEL #####

```{r}
graph_knn <- 
  po("encode", method = "one-hot") %>>%
  po("scale") %>>%
  lrn("regr.kknn")
```

```{r}
learner_knn <- as_learner(graph_knn)
```

```{r}
learner_knn$train(task, row_ids = train_set)
```

```{r}
prediction_knn <- learner_knn$predict(task, row_ids = test_set)
```

```{r}
prediction_knn$score(msr("regr.rmse"))
prediction_knn$score(msr("regr.mae"))
prediction_knn$score(msr("regr.rsq"))
```

#### KNN TUNING ####

```{r}
learner_knn <- lrn(
  "regr.kknn",
  k = to_tune(1, 50)
)
```

```{r}
graph_knn <- 
  po("encode", method = "one-hot") %>>%
  po("scale") %>>%
  learner_knn

graph_learner_knn <- as_learner(graph_knn)
```

```{r}
at_knn <- auto_tuner(
  tuner = tnr("random_search"),
  learner = graph_learner_knn,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 20)
)
```

```{r}
at_knn$train(task, row_ids = train_set)
```

```{r}
at_knn$tuning_result
```

```{r}
pred_knn <- at_knn$predict(task, row_ids = test_set)

pred_knn$score(msr("regr.rmse"))
pred_knn$score(msr("regr.mae"))
pred_knn$score(msr("regr.rsq"))
```

##### DEFAULT RF MODEL #####

```{r}
task_rf <- TaskRegr$new(
  id = "claim_rf",
  backend = MPL_data_lump,
  target = "ClaimAmount"
)
```

```{r}
learner_rf <- lrn(
  "regr.ranger",             #ranger implements Random Forest
  importance = "impurity",
  predict_type = "response"
)
```

```{r}
learner_rf$train(task, row_ids = train_set)
```

```{r}
pred_rf <- learner_rf$predict(task, row_ids = test_set)

pred_rf$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

#### TARGET ENCODING ######

```{r}
target_encoder <- po(
  "encodeimpact",
  affect_columns = selector_type(c("factor", "character"))
)
```

```{r}
learner_rf <- lrn(
  "regr.ranger",
  importance = "impurity",
  predict_type = "response"
)
```

```{r}
graph_rf_te <- target_encoder %>>% learner_rf

graph_learner_rf_te <- as_learner(graph_rf_te)
```

```{r}
graph_learner_rf_te$train(
  task,
  row_ids = train_set
)
```

```{r}
pred_rf_te <- graph_learner_rf_te$predict(
  task,
  row_ids = test_set
)
```

```{r}
pred_rf_te$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

The default Random Forest model performed better overall than the target encoded version. Although target encoding slightly improved the MAE, it led to a higher RMSE and a lower RSQ, suggesting that the model explained less of the variation in the data. Since Random Forest can already handle categorical variables quite well, target encoding did not provide any clear improvement in this case. Therefore, the default Random Forest model is chosen for further tuning.

##### TUNING ######

```{r}
search_space <- ps(
  mtry = p_int(lower = 2, upper = 12),
  min.node.size = p_int(lower = 1, upper = 20),
  sample.fraction = p_dbl(lower = 0.5, upper = 1),
  num.trees = p_int(lower = 500, upper = 1500)
)
```

```{r}
at_rf <- auto_tuner(
  tuner = tnr("random_search"),
  learner = learner_rf,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  search_space = search_space,
  terminator = trm("evals", n_evals = 25)
)
```

```{r}
at_rf$train(task, row_ids = train_set)
```

```{r}
at_rf$tuning_result
```

```{r}
pred_tuned <- at_rf$predict(task, row_ids = test_set)

pred_tuned$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

### RESULTS RF ###

```{r}
results_rf <- data.frame(
  Model = c("Baseline", "RF Default", "RF Target Encoded", "RF Tuned"),
  RMSE = c(pred_baseline$score(msr("regr.rmse")), pred_rf$score(msr("regr.rmse")), pred_rf_te$score(msr("regr.rmse")), pred_tuned$score(msr("regr.rmse"))),
  MAE = c(pred_baseline$score(msr("regr.mae")), pred_rf$score(msr("regr.mae")), pred_rf_te$score(msr("regr.mae")), pred_tuned$score(msr("regr.mae"))),
  RSQ = c(pred_baseline$score(msr("regr.rsq")), pred_rf$score(msr("regr.rsq")), pred_rf_te$score(msr("regr.rsq")), pred_tuned$score(msr("regr.rsq")))
)
results_rf
```

##### DEFAULT XGBoost MODEL #####

```{r}
learner_xgb_default <- as_learner(
  po("encode", method = "one-hot") %>>%
    lrn(
      "regr.xgboost",
      objective = "reg:squarederror",
      eval_metric = "rmse"
    )
)
```

```{r}
learner_xgb_default$train(
  task,
  row_ids = train_set
)
```

```{r}
pred_xgb_default <- learner_xgb_default$predict(
  task,
  row_ids = test_set
)
```

```{r}
pred_xgb_default$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

#### TARGET ENCODING #####

```{r}
learner_xgb_te <- as_learner(
  po("encodeimpact") %>>% #encodeimpact performs target encoding
    lrn(
      "regr.xgboost",
      objective = "reg:squarederror",
      eval_metric = "rmse"
    )
)
```

```{r}
learner_xgb_te$train(
  task,
  row_ids = train_set
)
```

```{r}
pred_xgb_te <- learner_xgb_te$predict(
  task,
  row_ids = test_set
)
```

```{r}
pred_xgb_te$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

The performance of the two models with the different encoding methods was very similar. However, the one-hot encoding lead to a slightly better model performance and therefore, we proceed with tuning that models hyperparameters.

##### TUNING #####

```{r}
learner_xgb_tuned <- as_learner(
  po("encode", method = "one-hot") %>>%
    lrn(
      "regr.xgboost",
      objective = "reg:squarederror",
      eval_metric = "rmse"
    )
)
```

```{r}
search_space <- ps(
  regr.xgboost.eta = p_dbl(0.01, 0.3),
  regr.xgboost.max_depth = p_int(3, 10),
  regr.xgboost.subsample = p_dbl(0.5, 1),
  regr.xgboost.colsample_bytree = p_dbl(0.5, 1),
  regr.xgboost.min_child_weight = p_int(1, 10),
  regr.xgboost.nrounds = p_int(100, 1000)
)
```

```{r}
tuner <- tnr("random_search")
```

```{r}
instance <- TuningInstanceBatchSingleCrit$new(
  task = task,
  learner = learner_xgb_tuned,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  search_space = search_space,
  terminator = trm("evals", n_evals = 50)
)
```

```{r}
plan(multisession)
tuner$optimize(instance)
```


```{r}
learner_xgb_tuned$param_set$values <- instance$result_learner_param_vals
```

```{r}
learner_xgb_tuned$train(task_xgb, row_ids = train_set)
```

```{r}
pred_xgb_tuned <- learner_xgb_tuned$predict(task, row_ids = test_set)

pred_xgb_tuned$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

###### RESULTS XGBoost ######

```{r}
results_xgb <- data.frame(
  Model = c("Baseline", "XGB Default", "XGB Target Encoded", "XGB Tuned"),
  RMSE = c(pred_baseline$score(msr("regr.rmse")), pred_xgb_default$score(msr("regr.rmse")), pred_xgb_te$score(msr("regr.rmse")), pred_xgb_tuned$score(msr("regr.rmse"))),
  MAE = c(pred_baseline$score(msr("regr.mae")), pred_xgb_default$score(msr("regr.mae")), pred_xgb_te$score(msr("regr.mae")), pred_xgb_tuned$score(msr("regr.mae"))),
  RSQ = c(pred_baseline$score(msr("regr.rsq")), pred_xgb_default$score(msr("regr.rsq")), pred_xgb_te$score(msr("regr.rsq")), pred_xgb_tuned$score(msr("regr.rsq")))
)
results_xgb
```

##### DEFAULT LIGHTGBM MODEL #####

```{r}
learner_lgb_default <- as_learner(
  po("encode", method = "one-hot") %>>%
    lrn("regr.lightgbm",
        objective = "regression",
        verbose = -1
    )
)
```

```{r}
learner_lgb_default$train(
  task,
  row_ids = train_set
)
```

```{r}
pred_lgb_default <- learner_lgb_default$predict(
  task,
  row_ids = test_set
)
```

```{r}
pred_lgb_default$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

#### TARGET ENCODING ####

```{r}
learner_lgb_te <- as_learner(
  po("encodeimpact") %>>%
    lrn(
      "regr.lightgbm",
      objective = "regression",
      verbose = -1
    )
)
```

```{r}
learner_lgb_te$train(
  task,
  row_ids = train_set
)
```

```{r}
pred_lgb_te <- learner_lgb_te$predict(
  task,
  row_ids = test_set
)
```

```{r}
pred_lgb_te$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

Once again, the target encoding does not improve performance, so we proceed with tuning the hyperparameters of the default model.

##### TUNING ####

```{r}
learner_lgb_tuned <- as_learner(
  po("encode", method = "one-hot") %>>%
    lrn("regr.lightgbm",
        objective = "regression",
        verbose = -1
    )
)
```

```{r}
search_space <- ps(
  regr.lightgbm.learning_rate =
    p_dbl(0.01, 0.1),
  regr.lightgbm.num_leaves =
    p_int(20, 80),
  regr.lightgbm.max_depth =
    p_int(3, 8),
  regr.lightgbm.feature_fraction =
    p_dbl(0.7, 1),
  regr.lightgbm.bagging_fraction =
    p_dbl(0.7, 1),
  regr.lightgbm.min_data_in_leaf =
    p_int(20, 80),
  regr.lightgbm.num_iterations =
    p_int(100, 500)
)
```

```{r}
tuner <- tnr("random_search")
```

```{r}
instance <- TuningInstanceBatchSingleCrit$new(
  task = task,
  learner = learner_lgb_tuned,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  search_space = search_space,
  terminator = trm("evals", n_evals = 50)
)
```

```{r}
plan(multisession)
tuner$optimize(instance)
```


```{r}
learner_lgb_tuned$param_set$values <- instance$result_learner_param_vals
```

```{r}
learner_lgb_tuned$train(task_lgb, row_ids = train_set)
```

```{r}
pred_lgb_tuned <- learner_lgb_tuned$predict(task_lgb, row_ids = test_set)

pred_lgb_tuned$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

###### RESULTS LGBM ######

```{r}
results_lgb <- data.frame(
  Model = c("Baseline", "LGBM Default", "LGBM Target Encoded", "LGBM Tuned"),
  RMSE = c(pred_baseline$score(msr("regr.rmse")), pred_lgb_default$score(msr("regr.rmse")), pred_lgb_te$score(msr("regr.rmse")), pred_lgb_tuned$score(msr("regr.rmse"))),
  MAE = c(pred_baseline$score(msr("regr.mae")), pred_lgb_default$score(msr("regr.mae")), pred_lgb_te$score(msr("regr.mae")), pred_lgb_tuned$score(msr("regr.mae"))),
  RSQ = c(pred_baseline$score(msr("regr.rsq")), pred_lgb_default$score(msr("regr.rsq")), pred_lgb_te$score(msr("regr.rsq")), pred_lgb_tuned$score(msr("regr.rsq")))
)
results_lgb
```

##### DEFAULT NEURAL NETWORK MODEL #####

```{r}
# MLP (Multi layer perceptron): Adam optimizer. SGD on mini batches (batch size = 32). See mlr_learners$get("regr.mlp")
graph_nn =
  po("encode", method = "one-hot") %>>%
  po("scale") %>>%
  lrn(
    "regr.mlp",
    epochs = 100,
    batch_size = 32,
    patience = 10, # Stopping process after 10 epochs without any improvement on validation set.
    measures_valid = msr("regr.rmse")
  )
```

```{r}
learner_nn = as_learner(graph_nn)
```

```{r}
set_validate(learner_nn, validate = 0.2) 
# Splitting the training data into new training and validation sets. This is used for early stopping.
```

```{r}
learner_nn$train(
  task,
  row_ids = train_set
)
```

```{r}
pred_nn = learner_nn$predict(
  task,
  row_ids = test_set
)
```

```{r}
pred_nn$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

###### RESULTS NN ######

```{r}
results_nn <- data.frame(
  Model = c("Baseline", "NN Default"),
  RMSE = c(pred_baseline$score(msr("regr.rmse")), pred_nn$score(msr("regr.rmse"))),
  MAE = c(pred_baseline$score(msr("regr.mae")), pred_nn$score(msr("regr.mae"))),
  RSQ = c(pred_baseline$score(msr("regr.rsq")), pred_nn$score(msr("regr.rsq")))
)
results_nn
```
