---
title: "Aflevering"
output: pdf_document
date: "2026-05-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Library
```{r}
library(mlr3)
library(mlr3learners)
library(mlr3tuning)
library(mlr3mbo)
library(mlr3pipelines)
library(paradox)
library(readr)
library(tidyverse)
library(mlr3viz)
```

# Load data

```{r}
freMPL <- read_csv("freMPL.csv")
```

# Pre-processing

We remove ClaimInd from the dataset, since we want to predict a price for a given customer, and we don't know, if the customer will have a claim.
Remove ...1 since it is just row_numbers.
```{r}
MPL_data <- subset(freMPL, select = -c(ClaimInd,...1))
```

We replace null-values from RecordEnd, so we don't have to deal with null values.
```{r}
MPL_data$RecordEnd[is.na(MPL_data$RecordEnd)] <- as.Date("2005-01-01")
```

Ordered factors and factors.
```{r}
# VehAge is an ordered variable
freMPL$VehAge <- factor(freMPL$VehAge, levels = c("0", "1", "2", "3", "4", "5", "6-7", "8-9", "10+"))

# VehMaxSpeed is an ordered variable
freMPL$VehMaxSpeed <- factor(freMPL$VehMaxSpeed,
                             levels = c("1-130 km/h", "130-140 km/h", "140-150 km/h",
                                        "150-160 km/h", "160-170 km/h", "170-180 km/h", 
                                        "180-190 km/h", "190-200 km/h", "200-220 km/h",
                                        "220+ km/h"))

# HasKmLimit
freMPL$HasKmLimit <- factor(freMPL$HasKmLimit)
```

Creating a new data set where few observations are lumped together as "other" to avoid problems with test-observations that are not not seen in the training data. 
```{r}
MPL_data_lump <- MPL_data

factor_cols <- names(MPL_data_lump)[sapply(MPL_data_lump, function(x) is.factor(x) || is.character(x))]

for (col in factor_cols) {
  MPL_data_lump[[col]] <- as.factor(MPL_data_lump[[col]])
  MPL_data_lump[[col]] <- fct_lump_min(MPL_data_lump[[col]], min = 10)
  MPL_data_lump[[col]] <- droplevels(MPL_data_lump[[col]])
}

categorical_vars <- c("VehAge", "Gender", "MariStat", "SocioCateg", "VehUsage", "VehBody", "VehPrice", "VehEngine", "VehEnergy", "VehMaxSpeed", "VehClass", "Garage") # VehMax Speed should perhaps be converted to a numeric variable by taking eg the max number in the observations given interval

MPL_data_lump[categorical_vars] <- lapply(MPL_data_lump[categorical_vars], factor)
```

Final data: 
```{r}
summary(MPL_data_lump)
```

# Task
```{r}
task = as_task_regr(MPL_data_lump, target = "ClaimAmount")
```

# Generisk kunde 1: 
Gennemsnits kunde
```{r}
gns_kunde <- MPL_data_lump[1,]

gns_kunde$Exposure <- 1  
gns_kunde$LicAge <- as.numeric(as.integer(mean(MPL_data_lump$LicAge)))
gns_kunde$RecordBeg <- as.Date("2004-01-01")
gns_kunde$RecordEnd <- as.Date("2005-01-01")
gns_kunde$VehAge <- factor(names(sort(table(MPL_data_lump$VehAge), decreasing = TRUE))[1],
                           levels = levels(MPL_data_lump$VehAge))
gns_kunde$Gender <- factor(names(sort(table(MPL_data_lump$Gender), decreasing = TRUE))[1],
                           levels = levels(MPL_data_lump$Gender))
gns_kunde$MariStat <- factor(names(sort(table(MPL_data_lump$MariStat), decreasing = TRUE))[1],
                             levels = levels(MPL_data_lump$MariStat))
gns_kunde$SocioCateg <- factor(names(sort(table(MPL_data_lump$SocioCateg), decreasing = TRUE))[1],
                               levels = levels(MPL_data_lump$SocioCateg))
gns_kunde$VehUsage <- factor(names(sort(table(MPL_data_lump$VehUsage), decreasing = TRUE))[1],
                             levels = levels(MPL_data_lump$VehUsage))
gns_kunde$DrivAge <- as.numeric(as.integer(mean(MPL_data_lump$DrivAge)))
gns_kunde$HasKmLimit <- as.numeric(names(sort(table(MPL_data_lump$HasKmLimit), decreasing = TRUE))[1])
gns_kunde$BonusMalus <- as.numeric(as.integer(mean(MPL_data_lump$BonusMalus)))
gns_kunde$VehBody <- factor(names(sort(table(MPL_data_lump$VehBody), decreasing = TRUE))[1],
                            levels = levels(MPL_data_lump$VehBody))
gns_kunde$VehPrice <- factor(names(sort(table(MPL_data_lump$VehPrice), decreasing = TRUE))[1],
                             levels = levels(MPL_data_lump$VehPrice))
gns_kunde$VehEngine <- factor(names(sort(table(MPL_data_lump$VehEngine), decreasing = TRUE))[1],
                              levels = levels(MPL_data_lump$VehEngine))
gns_kunde$VehEnergy <- factor(names(sort(table(MPL_data_lump$VehEnergy), decreasing = TRUE))[1],
                              levels = levels(MPL_data_lump$VehEnergy))
gns_kunde$VehMaxSpeed <- factor(names(sort(table(MPL_data_lump$VehMaxSpeed), decreasing = TRUE))[1],
                                levels = levels(MPL_data_lump$VehMaxSpeed))
gns_kunde$VehClass <- factor(names(sort(table(MPL_data_lump$VehClass), decreasing = TRUE))[1],
                             levels = levels(MPL_data_lump$VehClass))
gns_kunde$ClaimAmount <- mean(MPL_data_lump$ClaimAmount)
gns_kunde$RiskVar <- as.numeric(as.integer(mean(MPL_data_lump$RiskVar)))
gns_kunde$Garage <- factor(names(sort(table(MPL_data_lump$Garage), decreasing = TRUE))[1],
                           levels = levels(MPL_data_lump$Garage))
```

# Generisk kunde 2: 
Ekstrem kunde




