---
title: "Aflevering"
output: pdf_document
date: "2026-05-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Library
```{r}
library(mlr3)
library(mlr3learners)
library(mlr3tuning)
library(mlr3mbo)
library(mlr3pipelines)
library(paradox)
library(readr)
library(tidyverse)
library(mlr3viz)
```

# Load data
Below is data loaded, and ajustments are made: 
```{r}
freMPL <- read_csv("freMPL.csv")

# VehAge is an ordered variable, this ordering is taken into consideration
freMPL$VehAge <- factor(freMPL$VehAge, levels = c("0", "1", "2", "3", "4", "5", "6-7", "8-9", "10+"))

# VehMaxSpeed is an ordered variable, this ordering is taken into consideration
freMPL$VehMaxSpeed <- factor(freMPL$VehMaxSpeed,
                             levels = c("1-130 km/h", "130-140 km/h", "140-150 km/h",
                                        "150-160 km/h", "160-170 km/h", "170-180 km/h", 
                                        "180-190 km/h", "190-200 km/h", "200-220 km/h",
                                        "220+ km/h"))

freMPL <- freMPL[,-5]

# Do we need to do something to SocioCateg?
# Why do ClaimInt = 0 have negative ClaimAmount? subset = freMPL[freMPL$ClaimInd == 0,]
```

##Week 3
```{r}
categorical_vars <- c("RecordBeg","Gender","MariStat","SocioCateg","VehUsage","VehBody","VehPrice","VehEngine","VehEnergy","VehClass", "Garage")
freMPL$data[categorical_vars] <- lapply(freMPL$data[categorical_vars], factor)
```
```{r}
freMPL$RecordBeg <- factor(freMPL$RecordBeg)
freMPL$Gender <- factor(freMPL$Gender)
freMPL$MariStat <- factor(freMPL$MariStat)
freMPL$SocioCateg <- factor(freMPL$SocioCateg)
freMPL$VehUsage <- factor(freMPL$VehUsage)
freMPL$VehBody <- factor(freMPL$VehBody)
freMPL$VehPrice <- factor(freMPL$VehPrice)
freMPL$VehEngine <- factor(freMPL$VehEngine)
freMPL$VehEnergy <- factor(freMPL$VehEnergy)
freMPL$VehClass <- factor(freMPL$VehClass)
freMPL$Garage <- factor(freMPL$Garage)


```

We have to fit a linear model count ~ LicAge first.
```{r}
task_count_on_windspeed = as_task_regr(freMPL, target="ClaimInd")$select("LicAge")
lrn_count_on_windspeed = lrn("regr.lm")
lrn_count_on_windspeed$train(task_count_on_windspeed)
intercept = lrn_count_on_windspeed$model$coefficients[1]
univariate_windspeed_slope = lrn_count_on_windspeed$model$coefficients[2]

ggplot() +
  geom_point(mapping=aes(x = task_count_on_windspeed$data(cols=c("LicAge"))$LicAge, y=task_count_on_windspeed$truth())) +
  geom_abline(intercept=intercept, slope=univariate_windspeed_slope, color="red") +
  labs(x="LicAge", y="Count")
```
Now, let's fit a model using all coefficients.

```{r}
task_count_on_all = as_task_regr(freMPL, target="ClaimInd")
lrn_count_on_all = as_learner(po("encode", method="treatment") %>>% lrn("regr.lm"))
lrn_count_on_all$train(task_count_on_all)

joint_windspeed_slope = lrn_count_on_all$pipeops$regr.lm$learner_model$model$coefficients["LicAge"]

print(univariate_windspeed_slope)
print(joint_windspeed_slope)
```

Now, let's do the orthogonalization procedure from the FWL theorem.

```{r}
task_count_on_allother = as_task_regr(freMPL, target="ClaimInd")$select(setdiff(names(freMPL), c("LicAge", "ClaimInd")))
task_windspeed_on_allother = as_task_regr(freMPL, target="LicAge")$select(setdiff(names(freMPL), c("LicAge", "ClaimInd")))

lrn_count_on_allother = as_learner(po("encode", method="treatment") %>>% lrn("regr.lm"))
lrn_count_on_allother$train(task_count_on_allother)
pred_count_on_allother = lrn_count_on_allother$predict(task_count_on_allother)
count_residuals = pred_count_on_allother$truth - pred_count_on_allother$response

lrn_windspeed_on_allother = as_learner(po("encode", method="treatment") %>>% lrn("regr.lm"))
lrn_windspeed_on_allother$train(task_windspeed_on_allother)
pred_windspeed_on_allother = lrn_windspeed_on_allother$predict(task_windspeed_on_allother)
windspeed_residuals = pred_windspeed_on_allother$truth - pred_windspeed_on_allother$response

count_res_on_windspeed_res = lm(count_residuals ~ windspeed_residuals)
print(count_res_on_windspeed_res$coefficients[2])
print(joint_windspeed_slope)
```
We should also plot the FWL residual regression.

```{r}
ggplot() +
  geom_point(mapping=aes(x=windspeed_residuals, y=count_residuals)) +
  geom_abline(
    slope=count_res_on_windspeed_res$coefficients[2],
    intercept=count_res_on_windspeed_res$coefficients[1],
    color="red"
    )
```

Now, we have to repeat this procedure, but using cross-fitting and a non-linear nuisance regression.

We start by defining the tasks and learners.

```{r}
# task_count_on_allother # Already defined above
# task_windspeed_on_allother # Already defined above
lrn_count_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimInd ~ offset(log(Exposure)) + RecordBeg + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=20)
)
lrn_windspeed_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=LicAge ~ offset(log(Exposure)) + RecordBeg + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=20)
)


cross_fit = function(task, learner, K){
  set.seed(0)
  outer_resampling = rsmp("cv", folds=K)
  rr = resample(task, learner, outer_resampling)
  pred = rr$prediction() 
  pred$truth - pred$response
}

count_residuals_smooth = cross_fit(task_count_on_allother, lrn_count_on_allother_smooth, K=5)
windspeed_residuals_smooth = cross_fit(task_windspeed_on_allother, lrn_windspeed_on_allother_smooth, K=5)
```
After having performed the cross fitting, we fit the ordinary linear model for the coefficient for windspeed.

```{r}
lm_count_on_windspeed_smooth = lm(count_residuals_smooth ~ windspeed_residuals_smooth)
lm_count_on_windspeed_smooth$coefficients[2] # Slope
# Q: Is this supposed to be differnt from the previously estimated slope? Why?
```

```{r}
ggplot() + 
  geom_point(mapping = aes(x=windspeed_residuals_smooth, y=count_residuals_smooth)) +
  geom_abline(slope=lm_count_on_windspeed_smooth$coefficients[2], intercept=lm_count_on_windspeed_smooth$coefficients[1], color="red")
```

