---
title: "Aflevering"
output: pdf_document
date: "2026-05-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library
```{r}
library(mlr3)
library(mlr3learners)
library(mlr3tuning)
library(mlr3mbo)
library(mlr3pipelines)
library(paradox)
library(readr)
library(tidyverse)
library(mlr3viz)

library(rpart.plot) # Visualisering af decision trees
```

# Load data and overview
Below is data loaded, and ajustments are made: 
```{r}
freMPL <- read_csv("freMPL.csv")

# VehAge is an ordered variable, this ordering is taken into consideration
freMPL$VehAge <- factor(freMPL$VehAge, levels = c("0", "1", "2", "3", "4", "5", "6-7", "8-9", "10+"))

# VehMaxSpeed is an ordered variable, this ordering is taken into consideration
freMPL$VehMaxSpeed <- factor(freMPL$VehMaxSpeed,
                             levels = c("1-130 km/h", "130-140 km/h", "140-150 km/h",
                                        "150-160 km/h", "160-170 km/h", "170-180 km/h", 
                                        "180-190 km/h", "190-200 km/h", "200-220 km/h",
                                        "220+ km/h"))

# Do we need to do something to SocioCateg?
# Why do ClaimInt = 0 have negative ClaimAmount? subset = freMPL[freMPL$ClaimInd == 0,]

# Variable, der skal fjernes 
freMPL <- subset(freMPL, select = -`...1`) 
```

```{r}
View(freMPL)
summary(freMPL)
table(freMPL$VehAge)
table(freMPL$MariStat)
table(freMPL$SocioCateg)
```

```{r}
plot(freMPL$DrivAge, freMPL$LicAge)
plot(freMPL$DrivAge, freMPL$BonusMalus)
plot(freMPL$DrivAge, freMPL$ClaimInd)
```

```{r}
table(freMPL$VehBody, freMPL$VehUsage)
```
```{r}
names(freMPL)
```

An overview of all the covariates: 
```{r}
# Exposure
ggplot(freMPL, aes(x = Exposure)) + 
  geom_histogram(bins = 50, fill = "blue", col = "white") + 
  theme_minimal()
# LicAge
ggplot(freMPL, aes(x = LicAge/12)) + 
  geom_histogram(bins = 50, fill = "blue", col = "white") + 
  xlab("LicAge in years") + 
  theme_minimal()
# RecordBeg
ggplot(freMPL, aes(x = RecordBeg)) + 
  geom_histogram(bins = 13, fill = "blue", col = "white") + 
  theme_minimal()
# RecordEnd
ggplot(freMPL, aes(x = RecordEnd)) + 
  geom_histogram(bins = 13, fill = "blue", col = "white") + 
  theme_minimal()
# VehAge
ggplot(freMPL, aes(x = VehAge)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# Gender
ggplot(freMPL, aes(x = Gender)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# Maristat
ggplot(freMPL, aes(x = MariStat)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# SocioCateg
ggplot(freMPL, aes(x = SocioCateg)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
# VehUsage
ggplot(freMPL, aes(x = VehUsage)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# DrivAge
ggplot(freMPL, aes(x = DrivAge)) + 
  geom_histogram(bins = 30, fill = "blue", col = "white") + 
  theme_minimal()
# HasKmLimit (A numeric, 1 if there is a km limit for the policy, 0 otherwise)
ggplot(freMPL, aes(x = HasKmLimit)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# BonusMalus
ggplot(freMPL, aes(x = BonusMalus)) + 
  geom_histogram(bins = 30, fill = "blue", col = "white") + 
  theme_minimal()
# VehBody
ggplot(freMPL, aes(x = VehBody)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
# VehPrice (vehicle price from "A" (cheapest) to "Z" (most expensive))
ggplot(freMPL, aes(x = VehPrice)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# VehEngine
ggplot(freMPL, aes(x = VehEngine)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# VehEnergi
ggplot(freMPL, aes(x = VehEnergy)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# VehMaxSpeed
ggplot(freMPL, aes(x = VehMaxSpeed)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
# VehClass
ggplot(freMPL, aes(x = VehClass)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# RiskVar
ggplot(freMPL, aes(x = RiskVar)) + 
  geom_histogram(bins = 20, fill = "blue", col = "white") + 
  theme_minimal()
# Garage
ggplot(freMPL, aes(x = Garage)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
# ClaimInd (Claim indicator of the guarantee. (this is not the claim number))
ggplot(freMPL, aes(x = ClaimInd)) + 
  geom_bar(fill = "blue", col = "white") + 
  theme_minimal()
```

```{r}
range(freMPL$ClaimAmount)
freMPL[freMPL$ClaimAmount < 0,]
```

```{r}
plot(freMPL$Exposure,freMPL$ClaimAmount)
```


```{r}
# This below takes to long time: 
# subset_freMPL <- subset(freMPL, select = -c(SocioCateg, VehPrice)) # Too high cardinality
# sub_task = as_task_regr(subset_freMPL, target = "ClaimAmount")
# autoplot(sub_task, type = "pairs")
```

Responce variable: ClaimAmount
This variable looks heavytailed. 
```{r}
ggplot(freMPL, aes(x = ClaimAmount)) + 
  geom_histogram(bins = 20, fill = "blue", col = "white") + 
  theme_minimal()
```


# Noter: Hvilke ting tĂ¦nker vi at gennemgĂĄ?
Ridge regression, Lasso regression 
Spearman plot for at undersĂ¸ge korrelationer mellem features

# ForsĂ¸g pĂĄ at gentage det, vi lavede i Ă¸velserne i Week2_Sol:

## Decision tree

Define a task: 
```{r}
task = as_task_regr(freMPL, target = "ClaimAmount") # Regression task 
```

Put into train/test set
```{r}
set.seed(2026)
splits = partition(task, ratio = 0.8)

task_train = task$clone()$filter(splits$train)
task_test = task$clone()$filter(splits$test)
```

Define learner
```{r}
# my_learner = lrn("regr.rpart") # regression tree
# lrn("regr.rpart")$help() # Help
# Note: my_learner doesn't work because data contains characters, and rpart
# doesn't know how to handle this

task_train$feature_types

graph =
  po("colapply", applicator = as.factor, affect_columns = selector_type("character")) %>>%
  po("datefeatures") %>>%   # converts Date â†’ numeric features
  po("encode") %>>%
  lrn("regr.rpart")

learner = as_learner(graph)

learner$train(task_train)
learner$predict(task_test)$score(msrs(c("regr.mse","regr.mae")))
```
The above i basically a manual holdout. Holdout below: 
```{r}
set.seed(2026)

resampling = rsmp("holdout", ratio = 0.8)
rr = resample(task, learner, resampling)

rr$aggregate(msrs(c("regr.mse","regr.mae")))
```
Notice how this gives the same result as the manual approch above.

Take a look at the model
```{r}
learner$model

learner$model$regr.rpart$model
```
Comments on the regression tree: 
* First split is on ClaimInt: Makes sence to split between if a claim i covered (is this what ClaimInt tells? And does it makes sence to use this for predicting? We don't know for a future claim if it is rejected or not. But if it means wheather a claim is covered by the insurance, then it might make sence to keep it in the model)
* Does it logically makes sence to split by the day (seasons with split by month might make better sence)
* Does it make sence to split by day of week? 
* Split by BonusMalus around 100 makes sence (under 100 good, over bad)
* Split by LicAge makes sence: But it predicts lower for "young" customers?? (the once 
who havent had a licence for a long time) (but split by 571.9 ~ 47,7 years of divers licence 
--> makes driver older than 47,7+18 = 65,7)

```{r}
rpart.plot(learner$model$regr.rpart$model)
```


## Featureless learner: 
For comparison we train a featureless model (no covariates). 
This model just predicts the mean (no gain of infomation from covariates). 
```{r}
baseline <- lrn("regr.featureless")$train(task_train)
baseline$predict(task_test)$score(msrs(c("regr.mse","regr.mae")))
```
The regression tree performs better than featureless.
Note that data is pre-processed: 
* colapply fixes characters
* datefeatures fixes dates
* encode turns factors into numbers

Why use pipelines instead of cleaning data yourself: 
* The pre-processing is learned on the train-data
* The same pre-processing is used on the test-data (no "knowledge" from test-data
is leaked to the pre-processing part)
* Det skaber en fejl i resultat, hvis testdata benyttes til at konvertere data. 

Why use rsmp() and resample(): 
* Shorter code - the split is handled by the functions, and thus no need to split data.
* The result is the same though

# Linear regression
week 3

```{r}
# Create task - but removed some features because of missing values/not many values
task = as_task_regr(freMPL, target = "ClaimAmount") # Regression task -- same as above 
task$select(setdiff(task$feature_names, c("...1", "RecordEnd", "SocioCateg")))

# Define a learner
lrn_lm = as_learner(
    po("datefeatures") %>>% 
    #po("imputesample") %>>%
    po("encode", method="treatment") %>>% 
    lrn("regr.lm"))

# Train and test learner
set.seed(2026)
resampling = rsmp("holdout", ratio = 0.8)
rr = resample(task, lrn_lm, resampling)

rr$aggregate(msrs(c("regr.mse","regr.mae")))
```

What does po("encode", method="treatment") do: 
* creates a pipeline operator that converts categorical/factor variables into numeric dummy variables using treatment coding.
What does po("datefeatures") do: 
* converts dates into usable numeric variables such as year, month, day etc. 

# Clean comparison 
Define task
```{r}
task = as_task_regr(freMPL, target = "ClaimAmount")
task$select(setdiff(task$feature_names, c("...1", "RecordEnd", "SocioCateg")))
```

Mulige learners
```{r}
mlr_learners$keys("^regr")
```


Define learners for comparison 
```{r}
graph = po("colapply", applicator = as.factor, affect_columns = selector_type("character")) %>>%
  po("datefeatures") %>>%   # converts Date â†’ numeric features
  po("encode")

lrn_tree = as_learner(
  graph %>>%
  lrn("regr.rpart"))

lrn_lm = as_learner(
  graph %>>%
  lrn("regr.lm"))

lrn_featureless = as_learner(
  graph %>>%
  lrn("regr.featureless"))

lrn_gam = as_learner(
  graph %>>%
  lrn("regr.gam"))
```

Train and test-split to use
```{r}
set.seed(2026)
resampling = rsmp("holdout", ratio = 0.8)
```

```{r}
rr_tree = resample(task, lrn_tree, resampling)
rr_lm = resample(task, lrn_lm, resampling)
rr_featureless = resample(task, lrn_featureless, resampling)
rr_gam = resample(task, lrn_gam, resampling)
```

Score
```{r}
cat("tree model:\n")
rr_tree$aggregate(msrs(c("regr.mse", "regr.rmse", "regr.mae")))

cat("\nlm model:\n")
rr_lm$aggregate(msrs(c("regr.mse", "regr.rmse", "regr.mae")))

cat("\nfeatureless model:\n")
rr_featureless$aggregate(msrs(c("regr.mse", "regr.rmse", "regr.mae")))

cat("\ngam model:\n")
rr_gam$aggregate(msrs(c("regr.mse", "regr.rmse", "regr.mae")))
```

# GAM with auto-tuner 

```{r}
lrn_gam_tune = auto_tuner(
  tnr("random_search"),
  as_learner(
  graph %>>%
  lrn("regr.gam",
    #formula=count ~ s(temp) + s(feel_temp) + s(humidity) + season + year + month + hour + holiday + weekday + weather,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  )),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=20)
)

outer_resampling = rsmp("cv", folds=5)

rr_gam_tune = resample(task, lrn_gam_tune, outer_resampling)

rr_gam_tune$aggregate(msrs(c("regr.mse", "regr.rmse", "regr.mae")))
```

# Random forrest 

```{r}
lrn("regr.ranger")$help()
```

```{r}
task = as_task_regr(freMPL, target = "ClaimAmount") # Regression task 

graph =
  po("colapply", applicator = as.factor, affect_columns = selector_type("character")) %>>%
  po("datefeatures") %>>%   # converts Date â†’ numeric features
  po("encode") %>>%
  lrn("regr.rpart")

learner = as_learner(graph)

set.seed(2026)
resampling = rsmp("holdout", ratio = 0.8)

rr = resample(task, learner, resampling, store_models = TRUE) # True sĂĄ den kan hived ud efterfĂ¸lgende

rr$aggregate(msrs(c("regr.mse","regr.rmse","regr.mae")))
```

```{r}
rr$learners[[1]]$model$regr.rpart$model
```

PrĂ¸ver at kĂ¸re setup, hvor datoer fjernes, da det virker sjovt at selve ugedagen/tid pĂĄ mĂĄned mv. har indflydelse pĂĄ modellen 

```{r}
# Task 1 - fjerner slutdato
task = as_task_regr(freMPL, target = "ClaimAmount")$select(setdiff(names(freMPL), c("RecordEnd", "ClaimAmount"))) # Regression task uden datoer

# Task 2 - fjerner slutdato + claim index
task = as_task_regr(freMPL, target = "ClaimAmount")$select(setdiff(names(freMPL), c("RecordEnd", "ClaimAmount", "ClaimInd")))

# Task 3 - fjerner slutdato + startdato
task = as_task_regr(freMPL, target = "ClaimAmount")$select(setdiff(names(freMPL), c("RecordEnd", "ClaimAmount", "RecordBeg")))

graph =
  po("colapply", applicator = as.factor, affect_columns = selector_type("character")) %>>%
  po("datefeatures") %>>%   # converts Date â†’ numeric features
  po("encode") %>>%
  lrn("regr.rpart")

learner = as_learner(graph)

set.seed(2026)
resampling = rsmp("holdout", ratio = 0.8)

rr = resample(task, learner, resampling, store_models = TRUE) # True sĂĄ den kan hived ud efterfĂ¸lgende

rr$aggregate(msrs(c("regr.mse","regr.rmse","regr.mae")))

rr$learners[[1]]$model$regr.rpart$model
```
OvenstĂĄende task1 har en en bedre prediktion end den model, hvor der ikke pilles ved noget

task 0: Ingenting gjort
    regr.mse    regr.rmse     regr.mae 
987089.37673    993.52372     93.75692 

task 1: Enddate fjernet
    regr.mse    regr.rmse     regr.mae 
859213.36741    926.93763     90.42977 

task 2: Enddate + ClaimInd fjernet
   regr.mse   regr.rmse    regr.mae 
976935.7655    988.4006    185.8948 

task 3: Enddate + Startdate fjernet (bedre mse end task 1 - men dĂĄrligere mae)
    regr.mse    regr.rmse     regr.mae 
819371.73950    905.19155     91.16164 





