---
title: "IML_assignment"
author: "Philip Nordfang"
date: "2026-04-27"
output: html_document
---

```{r}
library(mlr3)
library(mlr3learners)
library(mlr3tuning)
library(mlr3mbo)
library(mlr3pipelines)
library(paradox)
library(future)
library(tidyverse)
library(forcats)
```


### Noter:

- GOAL: you should build models predicting ClaimAmount conditional on Exposure = 1
kagemand


```{r}
MPL_data = freMPL
```

```{r}
summary(MPL_data)
```

```{r}
data_exp <- subset(MPL_data, Exposure == 1)
```

```{r}
unique(data_exp$RecordEnd)
```

```{r}
hist(data_exp$ClaimAmount, main ="Claim amount distribution",
    xlab ="Claim Amount",
    col ="blue")
```

```{r}
data_exp <- subset(data_exp, select = -c(RecordEnd, RecordBeg, ClaimInd))
```

```{r}
# Creating a new data set where few observations are lumped together as "other" to avoid problems with test-observations that are not not seen in the training data
data_exp_lump <- data_exp

factor_cols <- names(data_exp_lump)[sapply(data_exp_lump, function(x) is.factor(x) || is.character(x))]

for (col in factor_cols) {
  data_exp_lump[[col]] <- as.factor(data_exp_lump[[col]])
  data_exp_lump[[col]] <- fct_lump_min(data_exp_lump[[col]], min = 10)
  data_exp_lump[[col]] <- droplevels(data_exp_lump[[col]])
}
```

```{r}
unique(data_exp_lump$SocioCateg)
```


#################################
####### Testing Exercise solutions from week 2 on the data

# Define task

```{r}
task = as_task_regr(data_exp_lump, target = "ClaimAmount")
```

# Create train and test set

```{r}
set.seed(2026)
splits = partition(task, ratio = 0.8)
task_train = task$clone()$filter(splits$train)
task_test = task$clone()$filter(splits$test)
```

#### define learner: linear regression

```{r}
my_lr_learner = lrn("regr.lm")
```


##### Creates a pipeline that converts categorical features to numeric with one hot encoding, standardizes numeric features, and applies linear regression.

```{r}
graph_learner_lr = as_learner(
  po("encode", method = "one-hot") %>>%
  po("scale") %>>%
  my_lr_learner
)
```


######## result on test set

```{r}
graph_learner_lr$train(task_train)
graph_learner_lr$predict(task_test)$score(msr("regr.rmse"))
```

### baseline with no features

```{r}
baseline <- lrn("regr.featureless")$train(task_train)
baseline$predict(task_test)$score(msr("regr.rmse"))
```

#################################
### Cross Validation

#### define learner

### elastic net
```{r}
my_elasticnet_learner = lrn(
  "regr.glmnet", 
  standardize = FALSE,
  alpha = to_tune(0, 1),
  lambda = to_tune(p_dbl(log(1e-5), log(1e1), trafo = exp))
  # Alternatively lambda = to_tune(1e-5, 1e1, logscale=TRUE)
  )
```

#### create a graph that first one hot-encodes factors, and afterwards applies elastic net
```{r}
graph_learner_elastic_net = as_learner(
  po("encode", method = "one-hot") %>>%
    po("scale") %>>%
    my_elasticnet_learner
)
```

```{r}
set.seed(2026)
instance = tune(
  tuner = tnr("random_search", batch_size=20), ### tuning method
  #tuner = tnr("mbo"), ### tuning method
  task = task_train,
  learner = graph_learner_elastic_net,
  resampling = rsmp("cv", folds = 5), #### resampling method: 5-fold cross validation
  measures = msr("regr.rmse"), #### root mean squared error
  terminator = trm("evals", n_evals = 50) #### terminator
)
```

```{r}
as.data.table(instance$archive$data)
instance$result_learner_param_vals
```

```{r}
instance$result_y
```

# keep an untuned copy of the same graph for the auto_tuner example below
```{r}
graph_learner_elastic_net_auto = graph_learner_elastic_net$clone(deep = TRUE)
```

######## result on test set

```{r}
graph_learner_elastic_net$param_set$values = instance$result_learner_param_vals
graph_learner_elastic_net$train(task_train)
graph_learner_elastic_net$predict(task_test)$score(msr("regr.rmse"))

archive_dt = as.data.table(instance$archive)
```


### sparsity / accuracy tradeoff for a few alpha values
```{r}
alpha_values = c(0, 0.25, 0.5, 0.75, 1)
lambda_grid = exp(seq(log(1e-5), log(1e1), length.out = 15))
tradeoff = expand.grid(alpha = alpha_values, lambda = lambda_grid)
tradeoff$n_features = NA_real_
tradeoff$cv_error = NA_real_
```


```{r}
for (i in seq_len(nrow(tradeoff))) {
  learner_i = as_learner(
    po("encode", method = "one-hot") %>>% po("scale") %>>%
      lrn("regr.glmnet",
          standardize = FALSE,
          alpha = tradeoff$alpha[i],
          lambda = tradeoff$lambda[i])
  )

  set.seed(2026)
  rr_i = resample(task_train, learner_i$clone(deep = TRUE), rsmp("cv", folds = 5))
  tradeoff$cv_error[i] = rr_i$aggregate(msr("regr.rmse"))

  learner_i$train(task_train)
  beta = learner_i$model$regr.glmnet$model$beta
  tradeoff$n_features[i] = sum(as.matrix(beta) != 0)
}
```

# Plot results using ggplot2
```{r}
ggplot(data=tradeoff) +
  geom_point(aes(x=n_features, y=cv_error, color=as.factor(alpha)))
```

# Plot results using base-r
```{r}
score_range = range(tradeoff$cv_error)
point_cols = c("#1a9850", "#91cf60", "#ffffbf", "#fc8d59", "#d73027")
line_cols = c("#1b9e77", "#66a61e", "#d95f02", "#e7298a", "#7570b3")

score_breaks = seq(score_range[1], score_range[2], length.out = length(point_cols) + 1)
color_id = cut(tradeoff$cv_error, breaks = score_breaks, include.lowest = TRUE, labels = FALSE)
tradeoff$point_col = point_cols[color_id]
legend_labels = paste0(
  sprintf("%.3f", score_breaks[-length(score_breaks)]),
  " - ",
  sprintf("%.3f", score_breaks[-1])
)


plot(NA,
     xlim = range(lambda_grid),
     ylim = range(tradeoff$n_features),
     log = "x",
     xlab = "lambda",
     ylab = "Number of non-zero coefficients",
     main = "Sparsity / accuracy tradeoff")

for (j in seq_along(alpha_values)) {
  alpha_j = alpha_values[j]
  tradeoff_j = tradeoff[tradeoff$alpha == alpha_j, ]

  lines(tradeoff_j$lambda, tradeoff_j$n_features, col = line_cols[j], lwd = 2)
  points(tradeoff_j$lambda, tradeoff_j$n_features,
         pch = 21,
         bg = tradeoff_j$point_col,
         col = line_cols[j],
         cex = 1.1)
}

legend("topright",
       legend = paste("alpha =", alpha_values),
       col = line_cols,
       lwd = 2,
       pch = 21,
       pt.bg = "white",
       bty = "n")

legend("bottomleft",
       legend = legend_labels,
       fill = point_cols,
       title = "CV error",
       bty = "n",
       cex = 0.8)
```


### alternative: use auto_tuner already here
```{r}
at_test = auto_tuner(
  tuner = tnr("random_search", batch_size = 20),
  learner = graph_learner_elastic_net_auto,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 50)
)

set.seed(2026)
at_test$train(task_train)
at_test$predict(task_test)$score(msr("regr.rmse"))
at_test$tuning_instance$result_learner_param_vals
```


### see beta
```{r}
graph_learner_elastic_net$model$regr.glmnet$model$beta
```



#################################
### Nested Cross Validation

```{r}
at = auto_tuner(
  tuner = tnr("random_search", batch_size=20),
  learner = graph_learner_elastic_net,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 50)
)

set.seed(2026)
rr = resample(task, at , rsmp("cv", folds = 5))
rr
rr$score(msr("regr.rmse"))[, .(iteration, regr.rmse)]
rr$aggregate(msr("regr.rmse"))

```

######## result on test set

```{r}
graph_learner_elastic_net$param_set$values = instance$result_learner_param_vals
graph_learner_elastic_net$train(task_train)
graph_learner_elastic_net$predict(task_test)$score(msr("regr.rmse"))

archive_dt = as.data.table(instance$archive)
```








