---
title: "IML_LightGBM"
author: "Philip Nordfang"
date: "2026-05-19"
output: html_document
---

```{r}
library(mlr3)
library(mlr3learners)
library(mlr3tuning)
library(mlr3mbo)
library(mlr3pipelines)
library(mlr3torch)
library(torch)
library(paradox)
library(future)
library(tidyverse)
library(forcats)
library(mlr3extralearners)
library(OpenML)
library(mgcv)
library(kknn)
library(lightgbm)
```

###### DATA HANDLING ######
```{r}
MPL_data = freMPL
```

```{r}
summary(MPL_data)
```


```{r}
hist(MPL_data$ClaimAmount, main ="Claim amount distribution",
    xlab ="Claim Amount",
    col ="blue")
```

```{r}
MPL_data <- subset(MPL_data, select = -ClaimInd)
MPL_data$RecordEnd <- as.Date(MPL_data$RecordEnd)
MPL_data$RecordEnd[is.na(MPL_data$RecordEnd)] <- as.Date("2005-01-01")
```

```{r}
# Creating a new data set where few observations are lumped together as "other" to avoid problems with test-observations that are not not seen in the training data
MPL_data_lump <- MPL_data

factor_cols <- names(MPL_data_lump)[sapply(MPL_data_lump, function(x) is.factor(x) || is.character(x))]

for (col in factor_cols) {
  MPL_data_lump[[col]] <- as.factor(MPL_data_lump[[col]])
  MPL_data_lump[[col]] <- fct_lump_min(MPL_data_lump[[col]], min = 10)
  MPL_data_lump[[col]] <- droplevels(MPL_data_lump[[col]])
}
```

```{r}
unique(MPL_data_lump$SocioCateg)
```

### convert categorical variables to factors
```{r}
categorical_vars <- c("RecordEnd", "VehAge", "Gender", "MariStat", "SocioCateg", "VehUsage", "VehBody", "VehPrice", "VehEngine", "VehEnergy", "VehMaxSpeed", "VehClass", "Garage") # VehMax Speed should perhaps be converted to a numeric variable by taking eg the max number in the observations given interval
MPL_data_lump[categorical_vars] <- lapply(MPL_data_lump[categorical_vars], factor)
```

##### DEFAULT LIGHTGBM MODEL #####

```{r}
task_lgb <- TaskRegr$new(
  id = "claim_amount",
  backend = MPL_data_lump,
  target = "ClaimAmount"
)
```

```{r}
set.seed(2026)

split <- partition(task_lgb, ratio = 0.8)

train_set <- split$train
test_set  <- split$test
```

```{r}
learner_lgb_default <- as_learner(
  po("encode", method = "one-hot") %>>%
    lrn("regr.lightgbm",
        objective = "regression",
        verbose = -1
    )
)
```

```{r}
learner_lgb_default$train(
  task_lgb,
  row_ids = train_set
)
```

```{r}
pred_lgb_default <- learner_lgb_default$predict(
  task_lgb,
  row_ids = test_set
)
```

```{r}
pred_lgb_default$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

#### TARGET ENCODING ####

```{r}
learner_lgb_te <- as_learner(
  po("encodeimpact") %>>%
    lrn(
      "regr.lightgbm",
      objective = "regression",
      verbose = -1
    )
)
```

```{r}
learner_lgb_te$train(
  task_lgb,
  row_ids = train_set
)
```

```{r}
pred_lgb_te <- learner_lgb_te$predict(
  task_lgb,
  row_ids = test_set
)
```

```{r}
pred_lgb_te$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

Once again, the target encoding does not improve performance, so we proceed with tuning the hyperparameters of the default model.

##### TUNING ####

```{r}
learner_lgb_tuned <- as_learner(
  po("encode", method = "one-hot") %>>%
    lrn("regr.lightgbm",
        objective = "regression",
        verbose = -1
    )
)
```

```{r}
search_space <- ps(
  regr.lightgbm.learning_rate =
    p_dbl(0.01, 0.1),
  regr.lightgbm.num_leaves =
    p_int(20, 80),
  regr.lightgbm.max_depth =
    p_int(3, 8),
  regr.lightgbm.feature_fraction =
    p_dbl(0.7, 1),
  regr.lightgbm.bagging_fraction =
    p_dbl(0.7, 1),
  regr.lightgbm.min_data_in_leaf =
    p_int(20, 80),
  regr.lightgbm.num_iterations =
    p_int(100, 500)
)
```

```{r}
tuner <- tnr("random_search")
```

```{r}
instance <- TuningInstanceBatchSingleCrit$new(
  task = task_lgb,
  learner = learner_lgb_tuned,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  search_space = search_space,
  terminator = trm("evals", n_evals = 50)
)
```

```{r}
plan(multisession)
tuner$optimize(instance)
```


```{r}
learner_lgb_tuned$param_set$values <- instance$result_learner_param_vals
```

```{r}
learner_lgb_tuned$train(task_lgb, row_ids = train_set)
```

```{r}
pred_lgb_tuned <- learner_lgb_tuned$predict(task_lgb, row_ids = test_set)

pred_lgb_tuned$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```


#### BASELINE MODEL FOR COMPARISON ######

```{r}
baseline <- lrn("regr.featureless")
```

```{r}
baseline$train(task_knn, row_ids = split_knn$train)
```

```{r}
pred_baseline <- baseline$predict(task_knn, row_ids = split_knn$test)
```

```{r}
pred_baseline$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```


###### RESULTS ######

```{r}
results_lgb <- data.frame(
  Model = c("Baseline", "LGBM Default", "LGBM Target Encoded", "LGBM Tuned"),
  RMSE = c(pred_baseline$score(msr("regr.rmse")), pred_lgb_default$score(msr("regr.rmse")), pred_lgb_te$score(msr("regr.rmse")), pred_lgb_tuned$score(msr("regr.rmse"))),
  MAE = c(pred_baseline$score(msr("regr.mae")), pred_lgb_default$score(msr("regr.mae")), pred_lgb_te$score(msr("regr.mae")), pred_lgb_tuned$score(msr("regr.mae"))),
  RSQ = c(pred_baseline$score(msr("regr.rsq")), pred_lgb_default$score(msr("regr.rsq")), pred_lgb_te$score(msr("regr.rsq")), pred_lgb_tuned$score(msr("regr.rsq")))
)
results_lgb
```
Hyperparameter tuning did not improve the predictive performance of the LightGBM model, suggesting that the default parameter settings were already sufficiently well adapted to the dataset.



