---
title: "IML_RF"
author: "Philip Nordfang"
date: "2026-05-18"
output: html_document
---

```{r}
library(mlr3)
library(mlr3learners)
library(mlr3tuning)
library(mlr3mbo)
library(mlr3pipelines)
library(mlr3torch)
library(torch)
library(paradox)
library(future)
library(tidyverse)
library(forcats)
library(mlr3extralearners)
library(OpenML)
library(mgcv)
library(kknn)
library(lightgbm)
```

###### DATA HANDLING ######
```{r}
MPL_data = freMPL
```

```{r}
summary(MPL_data)
```


```{r}
hist(MPL_data$ClaimAmount, main ="Claim amount distribution",
    xlab ="Claim Amount",
    col ="blue")
```

```{r}
MPL_data <- subset(MPL_data, select = -ClaimInd)
MPL_data$RecordEnd <- as.Date(MPL_data$RecordEnd)
MPL_data$RecordEnd[is.na(MPL_data$RecordEnd)] <- as.Date("2005-01-01")
```

```{r}
# Creating a new data set where few observations are lumped together as "other" to avoid problems with test-observations that are not not seen in the training data
MPL_data_lump <- MPL_data

factor_cols <- names(MPL_data_lump)[sapply(MPL_data_lump, function(x) is.factor(x) || is.character(x))]

for (col in factor_cols) {
  MPL_data_lump[[col]] <- as.factor(MPL_data_lump[[col]])
  MPL_data_lump[[col]] <- fct_lump_min(MPL_data_lump[[col]], min = 10)
  MPL_data_lump[[col]] <- droplevels(MPL_data_lump[[col]])
}
```

```{r}
unique(MPL_data_lump$SocioCateg)
```

### convert categorical variables to factors
```{r}
categorical_vars <- c("RecordEnd", "VehAge", "Gender", "MariStat", "SocioCateg", "VehUsage", "VehBody", "VehPrice", "VehEngine", "VehEnergy", "VehMaxSpeed", "VehClass", "Garage") # VehMax Speed should perhaps be converted to a numeric variable by taking eg the max number in the observations given interval
MPL_data_lump[categorical_vars] <- lapply(MPL_data_lump[categorical_vars], factor)
```

##### DEFAULT RF MODEL #####

```{r}
task_rf <- TaskRegr$new(
  id = "claim_rf",
  backend = MPL_data_lump,
  target = "ClaimAmount"
)
```

```{r}
set.seed(2026)

split <- partition(task_rf, ratio = 0.8)

train_set <- split$train
test_set  <- split$test
```

```{r}
learner_rf <- lrn(
  "regr.ranger",             #ranger implements Random Forest
  importance = "impurity",
  predict_type = "response"
)
```

```{r}
set.seed(2026)

learner_rf$train(task_rf, row_ids = train_set)
```

```{r}
pred_rf <- learner_rf$predict(task_rf, row_ids = test_set)

pred_rf$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

#### TARGET ENCODING ######

```{r}
target_encoder <- po(
  "encodeimpact",
  affect_columns = selector_type(c("factor", "character"))
)
```

```{r}
learner_rf <- lrn(
  "regr.ranger",
  importance = "impurity",
  predict_type = "response"
)
```

```{r}
graph_rf_te <- target_encoder %>>% learner_rf

graph_learner_rf_te <- as_learner(graph_rf_te)
```

```{r}
set.seed(2026)

graph_learner_rf_te$train(
  task_rf,
  row_ids = train_set
)
```

```{r}
pred_rf_te <- graph_learner_rf_te$predict(
  task_rf,
  row_ids = test_set
)
```

```{r}
pred_rf_te$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```

The default Random Forest model performed better overall than the target encoded version. Although target encoding slightly improved the MAE, it led to a higher RMSE and a lower RSQ, suggesting that the model explained less of the variation in the data. Since Random Forest can already handle categorical variables quite well, target encoding did not provide any clear improvement in this case. Therefore, the default Random Forest model is chosen for further tuning.

##### TUNING ######

```{r}
search_space <- ps(
  mtry = p_int(lower = 2, upper = 12),
  min.node.size = p_int(lower = 1, upper = 20),
  sample.fraction = p_dbl(lower = 0.5, upper = 1),
  num.trees = p_int(lower = 500, upper = 1500)
)
```

```{r}
at_rf <- auto_tuner(
  tuner = tnr("random_search"),
  learner = learner_rf,
  resampling = rsmp("cv", folds = 5),
  measure = msr("regr.rmse"),
  search_space = search_space,
  terminator = trm("evals", n_evals = 25)
)
```

```{r}
set.seed(2026)

at_rf$train(task_rf, row_ids = train_set)
```

```{r}
at_rf$tuning_result
```

```{r}
pred_tuned <- at_rf$predict(task_rf, row_ids = test_set)

pred_tuned$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```


#### BASELINE MODEL FOR COMPARISON ######

```{r}
baseline <- lrn("regr.featureless")
```

```{r}
baseline$train(task_knn, row_ids = split_knn$train)
```

```{r}
pred_baseline <- baseline$predict(task_knn, row_ids = split_knn$test)
```

```{r}
pred_baseline$score(list(
  msr("regr.rmse"),
  msr("regr.mae"),
  msr("regr.rsq")
))
```


###### RESULTS ######

```{r}
results_rf <- data.frame(
  Model = c("Baseline", "RF Default", "RF Target Encoded", "RF Tuned"),
  RMSE = c(pred_baseline$score(msr("regr.rmse")), pred_rf$score(msr("regr.rmse")), pred_rf_te$score(msr("regr.rmse")), pred_tuned$score(msr("regr.rmse"))),
  MAE = c(pred_baseline$score(msr("regr.mae")), pred_rf$score(msr("regr.mae")), pred_rf_te$score(msr("regr.mae")), pred_tuned$score(msr("regr.mae"))),
  RSQ = c(pred_baseline$score(msr("regr.rsq")), pred_rf$score(msr("regr.rsq")), pred_rf_te$score(msr("regr.rsq")), pred_tuned$score(msr("regr.rsq")))
)
results_rf
```

The tuned Random Forest model achieved the best overall performance among the tested Random Forest models. Compared to the default model, tuning reduced both RMSE and MAE while also improving the RSQ.




