---
title: "Partial linear regression"
author: "Ingelise Reffeldt"
date: "2026-05-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library
```{r}
library(mlr3)
library(mlr3learners)
library(mlr3extralearners)
library(mlr3tuning)
library(mlr3mbo)
library(mlr3pipelines)
library(paradox)
library(readr)
library(tidyverse)
library(mlr3viz)
library(Metrics)
```

We start by making a matrix for the results we want 
```{r}
results <- matrix(0, nrow=20, ncol=3)
rownames(results) = c("LicAge", "Exposure", "RecordBeg","RecordEnd","VehAge","Gender","MariStat","SocioCateg","VehUsage","DrivAge","HasKmLimit","BonusMalus","VehBody","VehPrice","VehEngine","VehEnergy","VehMaxSpeed","VehClass","RiskVar","Garage")
colnames(results) = c("regr.rmse", "regr.mae", "regr.rsq")
```

Now we make a model for each coefficient, where we focus on one off the coefficients and put all of the other together.

## LicAge
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("LicAge", "ClaimAmount")))
task_LicAge_on_allother = as_task_regr(MPL_data_lump, target="LicAge")$select(setdiff(names(freMPL), c("LicAge", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_LicAge_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=LicAge ~ Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

cross_fit = function(task, learner, K){
  set.seed(2026)
  outer_resampling = rsmp("cv", folds=K)
  rr = resample(task, learner, outer_resampling)
  pred = rr$prediction() 
  pred$truth - pred$response
}

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
LicAge_residuals_smooth = cross_fit(task_LicAge_on_allother, lrn_LicAge_on_allother_smooth, K=2)

lm_amount_on_LicAge_smooth = lm(amount_residuals_smooth ~ LicAge_residuals_smooth)

results[1,] <- c(sqrt(mean(lm_amount_on_LicAge_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_LicAge_smooth)),
summary(lm_amount_on_LicAge_smooth)$r.squared)
```
##Exposure 
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("Exposure", "ClaimAmount")))
task_Exposure_on_allother = as_task_regr(MPL_data_lump, target="Exposure")$select(setdiff(names(freMPL), c("Exposure", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_Exposure_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=Exposure ~ LicAge + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
Exposure_residuals_smooth = cross_fit(task_Exposure_on_allother, lrn_Exposure_on_allother_smooth, K=2)

lm_amount_on_Exposure_smooth = lm(amount_residuals_smooth ~ Exposure_residuals_smooth)

results[2,] <- c(sqrt(mean(lm_amount_on_Exposure_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_Exposure_smooth)),
summary(lm_amount_on_Exposure_smooth)$r.squared)
```
##RecordBeg 
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("RecordBeg", "ClaimAmount")))
task_RecordBeg_on_allother = as_task_regr(MPL_data_lump, target="RecordBeg")$select(setdiff(names(freMPL), c("RecordBeg", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_RecordBeg_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=RecordBeg ~ LicAge + Exposure + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
RecordBeg_residuals_smooth = cross_fit(task_RecordBeg_on_allother, lrn_RecordBeg_on_allother_smooth, K=2)

lm_amount_on_RecordBeg_smooth = lm(amount_residuals_smooth ~ RecordBeg_residuals_smooth)

results[3,] <- c(sqrt(mean(lm_amount_on_RecordBeg_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_RecordBeg_smooth)),
summary(lm_amount_on_RecordBeg_smooth)$r.squared)
```
##RecordEnd
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("RecordEnd", "ClaimAmount")))
task_RecordEnd_on_allother = as_task_regr(MPL_data_lump, target="RecordEnd")$select(setdiff(names(freMPL), c("RecordEnd", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_RecordEnd_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=RecordEnd ~ LicAge + Exposure + RecordBeg + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
RecordEnd_residuals_smooth = cross_fit(task_RecordEnd_on_allother, lrn_RecordEnd_on_allother_smooth, K=2)

lm_amount_on_RecordEnd_smooth = lm(amount_residuals_smooth ~ RecordEnd_residuals_smooth)

results[4,] <- c(sqrt(mean(lm_amount_on_RecordEnd_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_RecordEnd_smooth)),
summary(lm_amount_on_RecordEnd_smooth)$r.squared)
```
##VehAge
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("VehAge", "ClaimAmount")))
task_VehAge_on_allother = as_task_regr(MPL_data_lump, target="VehAge")$select(setdiff(names(freMPL), c("VehAge", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_VehAge_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=VehAge ~ LicAge + Exposure + RecordBeg + RecordEnd + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
VehAge_residuals_smooth = cross_fit(task_VehAge_on_allother, lrn_VehAge_on_allother_smooth, K=2)

lm_amount_on_VehAge_smooth = lm(amount_residuals_smooth ~ VehAge_residuals_smooth)

results[5,] <- c(sqrt(mean(lm_amount_on_VehAge_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_VehAge_smooth)),
summary(lm_amount_on_VehAge_smooth)$r.squared)
```
##Gender 
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("Gender", "ClaimAmount")))
task_Gender_on_allother = as_task_regr(MPL_data_lump, target="Gender")$select(setdiff(names(freMPL), c("Gender", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_Gender_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=Gender ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
Gender_residuals_smooth = cross_fit(task_Gender_on_allother, lrn_Gender_on_allother_smooth, K=2)

lm_amount_on_Gender_smooth = lm(amount_residuals_smooth ~ Gender_residuals_smooth)

results[6,] <- c(sqrt(mean(lm_amount_on_Gender_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_Gender_smooth)),
summary(lm_amount_on_Gender_smooth)$r.squared)
```
##MariStat 
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("MariStat", "ClaimAmount")))
task_MariStat_on_allother = as_task_regr(MPL_data_lump, target="MariStat")$select(setdiff(names(freMPL), c("MariStat", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_MariStat_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=MariStat ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
MariStat_residuals_smooth = cross_fit(task_MariStat_on_allother, lrn_MariStat_on_allother_smooth, K=2)

lm_amount_on_MariStat_smooth = lm(amount_residuals_smooth ~ MariStat_residuals_smooth)

results[7,] <- c(sqrt(mean(lm_amount_on_MariStat_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_MariStat_smooth)),
summary(lm_amount_on_MariStat_smooth)$r.squared)
```
##SocioCateg
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("SocioCateg", "ClaimAmount")))
task_SocioCateg_on_allother = as_task_regr(MPL_data_lump, target="SocioCateg")$select(setdiff(names(freMPL), c("SocioCateg", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_SocioCateg_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=SocioCateg ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
SocioCateg_residuals_smooth = cross_fit(task_SocioCateg_on_allother, lrn_SocioCateg_on_allother_smooth, K=2)

lm_amount_on_SocioCateg_smooth = lm(amount_residuals_smooth ~ SocioCateg_residuals_smooth)

results[8,] <- c(sqrt(mean(lm_amount_on_SocioCateg_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_SocioCateg_smooth)),
summary(lm_amount_on_SocioCateg_smooth)$r.squared)
```
##VehUsage
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("VehUsage", "ClaimAmount")))
task_VehUsage_on_allother = as_task_regr(MPL_data_lump, target="VehUsage")$select(setdiff(names(freMPL), c("VehUsage", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_VehUsage_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=VehUsage ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
VehUsage_residuals_smooth = cross_fit(task_VehUsage_on_allother, lrn_VehUsage_on_allother_smooth, K=2)

lm_amount_on_VehUsage_smooth = lm(amount_residuals_smooth ~ VehUsage_residuals_smooth)

results[9,] <- c(sqrt(mean(lm_amount_on_VehUsage_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_VehUsage_smooth)),
summary(lm_amount_on_VehUsage_smooth)$r.squared)
```
##DrivAge
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("DrivAge", "ClaimAmount")))
task_DrivAge_on_allother = as_task_regr(MPL_data_lump, target="DrivAge")$select(setdiff(names(freMPL), c("DrivAge", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_DrivAge_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=DrivAge ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + HasKmLimit + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
DrivAge_residuals_smooth = cross_fit(task_DrivAge_on_allother, lrn_DrivAge_on_allother_smooth, K=2)

lm_amount_on_DrivAge_smooth = lm(amount_residuals_smooth ~ DrivAge_residuals_smooth)

results[10,] <- c(sqrt(mean(lm_amount_on_DrivAge_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_DrivAge_smooth)),
summary(lm_amount_on_DrivAge_smooth)$r.squared)
```
##HasKmLimit 
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("HasKmLimit", "ClaimAmount")))
task_HasKmLimit_on_allother = as_task_regr(MPL_data_lump, target="HasKmLimit")$select(setdiff(names(freMPL), c("HasKmLimit", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_HasKmLimit_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=HasKmLimit ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + s(BonusMalus) + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
HasKmLimit_residuals_smooth = cross_fit(task_HasKmLimit_on_allother, lrn_HasKmLimit_on_allother_smooth, K=2)

lm_amount_on_HasKmLimit_smooth = lm(amount_residuals_smooth ~ HasKmLimit_residuals_smooth)

results[11,] <- c(sqrt(mean(lm_amount_on_HasKmLimit_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_HasKmLimit_smooth)),
summary(lm_amount_on_HasKmLimit_smooth)$r.squared)
```
##BonusMalus
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("BonusMalus", "ClaimAmount")))
task_BonusMalus_on_allother = as_task_regr(MPL_data_lump, target="BonusMalus")$select(setdiff(names(freMPL), c("BonusMalus", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_BonusMalus_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=BonusMalus ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + VehBody + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
BonusMalus_residuals_smooth = cross_fit(task_BonusMalus_on_allother, lrn_BonusMalus_on_allother_smooth, K=2)

lm_amount_on_BonusMalus_smooth = lm(amount_residuals_smooth ~ BonusMalus_residuals_smooth)

results[12,] <- c(sqrt(mean(lm_amount_on_BonusMalus_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_BonusMalus_smooth)),
summary(lm_amount_on_BonusMalus_smooth)$r.squared)
```
##VehBody
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("VehBody", "ClaimAmount")))
task_VehBody_on_allother = as_task_regr(MPL_data_lump, target="VehBody")$select(setdiff(names(freMPL), c("VehBody", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_VehBody_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=VehBody ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehPrice + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
VehBody_residuals_smooth = cross_fit(task_VehBody_on_allother, lrn_VehBody_on_allother_smooth, K=2)

lm_amount_on_VehBody_smooth = lm(amount_residuals_smooth ~ VehBody_residuals_smooth)

results[13,] <- c(sqrt(mean(lm_amount_on_VehBody_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_VehBody_smooth)),
summary(lm_amount_on_VehBody_smooth)$r.squared)
```
##VehPrice 
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("VehPrice", "ClaimAmount")))
task_VehPrice_on_allother = as_task_regr(MPL_data_lump, target="VehPrice")$select(setdiff(names(freMPL), c("VehPrice", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_VehPrice_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=VehPrice ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehEngine+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
VehPrice_residuals_smooth = cross_fit(task_VehPrice_on_allother, lrn_VehPrice_on_allother_smooth, K=2)

lm_amount_on_VehPrice_smooth = lm(amount_residuals_smooth ~ VehPrice_residuals_smooth)

results[14,] <- c(sqrt(mean(lm_amount_on_VehPrice_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_VehPrice_smooth)),
summary(lm_amount_on_VehPrice_smooth)$r.squared)
```
##VehEngine
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("VehEngine", "ClaimAmount")))
task_VehEngine_on_allother = as_task_regr(MPL_data_lump, target="VehEngine")$select(setdiff(names(freMPL), c("VehEngine", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VePrice+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_VehEngine_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=VehEngine ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice+VehEnergy+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
VehEngine_residuals_smooth = cross_fit(task_VehEngine_on_allother, lrn_VehEngine_on_allother_smooth, K=2)

lm_amount_on_VehEngine_smooth = lm(amount_residuals_smooth ~ VehEngine_residuals_smooth)

results[15,] <- c(sqrt(mean(lm_amount_on_VehEngine_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_VehEngine_smooth)),
summary(lm_amount_on_VehEngine_smooth)$r.squared)
```
##VehEnergy
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("VehEnergy", "ClaimAmount")))
task_VehEnergy_on_allother = as_task_regr(MPL_data_lump, target="VehEnergy")$select(setdiff(names(freMPL), c("VehEnergy", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VePrice+VehEngine+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_VehEnergy_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=VehEnergy ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice+VehEngine+VehMaxSpeed + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
VehEnergy_residuals_smooth = cross_fit(task_VehEnergy_on_allother, lrn_VehEnergy_on_allother_smooth, K=2)

lm_amount_on_VehEnergy_smooth = lm(amount_residuals_smooth ~ VehEnergy_residuals_smooth)

results[16,] <- c(sqrt(mean(lm_amount_on_VehEnergy_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_VehEnergy_smooth)),
summary(lm_amount_on_VehEnergy_smooth)$r.squared)
```
##VehMaxSpeed
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("VehMaxSpeed", "ClaimAmount")))
task_VehMaxSpeed_on_allother = as_task_regr(MPL_data_lump, target="VehMaxSpeed")$select(setdiff(names(freMPL), c("VehMaxSpeed", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VePrice+VehEngine+VehEnergy + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_VehMaxSpeed_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=VehMaxSpeed ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice+VehEngine+VehEnergy + VehClass + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
VehMaxSpeed_residuals_smooth = cross_fit(task_VehMaxSpeed_on_allother, lrn_VehMaxSpeed_on_allother_smooth, K=2)

lm_amount_on_VehMaxSpeed_smooth = lm(amount_residuals_smooth ~ VehMaxSpeed_residuals_smooth)

results[17,] <- c(sqrt(mean(lm_amount_on_VehMaxSpeed_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_VehMaxSpeed_smooth)),
summary(lm_amount_on_VehMaxSpeed_smooth)$r.squared)
```
##VehClass 
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("VehClass", "ClaimAmount")))
task_VehClass_on_allother = as_task_regr(MPL_data_lump, target="VehClass")$select(setdiff(names(freMPL), c("VehClass", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VePrice+VehEngine+VehEnergy + VehMaxSpeed + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_VehClass_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=VehClass ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice+VehEngine+VehEnergy + VehMaxSpeed + s(RiskVar) + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
VehClass_residuals_smooth = cross_fit(task_VehClass_on_allother, lrn_VehClass_on_allother_smooth, K=2)

lm_amount_on_VehClass_smooth = lm(amount_residuals_smooth ~ VehClass_residuals_smooth)

results[18,] <- c(sqrt(mean(lm_amount_on_VehClass_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_VehClass_smooth)),
summary(lm_amount_on_VehClass_smooth)$r.squared)
```
##RiskVar 
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("RiskVar", "ClaimAmount")))
task_RiskVar_on_allother = as_task_regr(MPL_data_lump, target="RiskVar")$select(setdiff(names(freMPL), c("RiskVar", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VePrice+VehEngine+VehEnergy + VehMaxSpeed + VehClass + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_RiskVar_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=RiskVar ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice+VehEngine+VehEnergy + VehMaxSpeed + VehClass + Garage,
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
RiskVar_residuals_smooth = cross_fit(task_RiskVar_on_allother, lrn_RiskVar_on_allother_smooth, K=2)

lm_amount_on_RiskVar_smooth = lm(amount_residuals_smooth ~ RiskVar_residuals_smooth)

results[19,] <- c(sqrt(mean(lm_amount_on_RiskVar_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_RiskVar_smooth)),
summary(lm_amount_on_RiskVar_smooth)$r.squared)
```
##Garage
```{r}
task_amount_on_allother = as_task_regr(MPL_data_lump, target="ClaimAmount")$select(setdiff(names(freMPL), c("Garage", "ClaimAmount")))
task_Garage_on_allother = as_task_regr(MPL_data_lump, target="Garage")$select(setdiff(names(freMPL), c("Garage", "ClaimAmount")))

lrn_amount_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=ClaimAmount ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VePrice+VehEngine+VehEnergy + VehMaxSpeed + VehClass + s(RiskVar),
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)
lrn_Garage_on_allother_smooth = auto_tuner(
  tnr("random_search"),
  lrn(
    "regr.gam",
    formula=Garage ~ LicAge + Exposure + RecordBeg + RecordEnd + VehAge + Gender + MariStat + SocioCateg + VehUsage + s(DrivAge) + HasKmLimit + s(BonusMalus) + VehBody + VehPrice+VehEngine+VehEnergy + VehMaxSpeed + VehClass + s(RiskVar),
    gamma = to_tune(1,1e3,logscale=TRUE),
    select = to_tune(c(TRUE, FALSE)),
    drop.unused.levels=FALSE # Default is TRUE, but this may cause errors if splits are unlucky
  ),
  measure = msr("regr.mse"),
  resampling = rsmp("cv", folds=2),
  terminator = trm("evals", n_evals=10)
)

amount_residuals_smooth = cross_fit(task_amount_on_allother, lrn_amount_on_allother_smooth, K=2)
Garage_residuals_smooth = cross_fit(task_Garage_on_allother, lrn_Garage_on_allother_smooth, K=2)

lm_amount_on_Garage_smooth = lm(amount_residuals_smooth ~ Garage_residuals_smooth)

results[20,] <- c(sqrt(mean(lm_amount_on_Garage_smooth$residuals^2)),
mae(MPL_data_lump$ClaimAmount, predict(lm_amount_on_Garage_smooth)),
summary(lm_amount_on_Garage_smooth)$r.squared)
```